% !TEX root = ../tocl-version.tex

\section{Introduction}
   \label{sec:intro}

\subsection{Expressiveness modulo bisimilarity}

A seminal result in the theory of modal logic is van Benthem's Characterisation
Theorem~\cite{vanBenthemPhD}, stating that, over  the class of all labelled transition systems (LTSs for short),  every bisimulation-invariant
first-order formula is equivalent to (the standard
translation of) a modal formula:
\begin{equation}
\label{eq:vB}
\ML \equiv \fo/{\bis}  \qquad  \text{ (over the class of all LTSs)}.
\end{equation}
Over the years, a wealth of variations of the Characterisation Theorem have been
obtained.
For instance, van Benthem's theorem is one of the few
preservation results that transfers to the setting of finite
models~\cite{rose:moda97}; for a recent, rich source of van Benthem-style
characterisation results, see~\cite{DawarO09}. 
The general pattern of these results takes the shape
\begin{equation}
\label{eq:vBGeneral}
M \equiv \yvlang/{\bis}  \qquad  
\text{ (over a class of models $\mathsf{C}$)}.
\end{equation}
Apart from their obvious relevance to model theory, the interest in these 
results increases if $\mathsf{C}$ consists of transition structures that 
represent certain computational processes, as in the theory of the formal 
specification and verification of properties of software.
In this context, one often takes the point of view that bisimilar models 
represent \emph{the same} process. 
For this reason, only bisimulation-invariant properties are relevant.
Seen in this light, \eqref{eq:vBGeneral} is an \emph{expressive completeness} 
result: all the relevant properties expressible in $\yvlang$ (which is generally
some rich yardstick formalism), can already be expressed in a (usually
computationally more feasible) modal fragment $M$.

Of special interest to us is the work~\cite{Jan96}, which extends van Benthem's
result to the setting of \emph{second-order} logic, by proving that the 
bisimulation-invariant fragment of standard monadic second-order logic 
($\smso$) is the \emph{modal $\mu$-calculus} ($\muML$), viz., the extension of
basic modal logic with least- and greatest fixpoint operators:
\begin{equation}
\label{eq:JW}
\muML \equiv \smso/{\bis}  \qquad  \text{ (over the class of all LTSs)}.
\end{equation}
The aim of this paper is to study the fine structure of such connections between
second-order logics and modal $\mu$-calculi, obtaining variations of the
expressiveness completeness results \eqref{eq:vB} and \eqref{eq:JW}.

Our departure point is a result from \cite{ArnNiw92} for the class of 
\emph{binary} trees.
Their setting is somewhat different from ours; in particular, since their trees
have explicit functions accessing the left- and right successor of a node,
the notion of bisimilarity trivialises to the isomorphism relation.
Nevertheless, the key observation of Arnold \& Niwi{\'n}nski is to link the 
\emph{alternation-free fragment} of a modal $\mu$-calculus to so-called 
\emph{weak} $\mso$, the version of monadic second-order logic where second-order
quantification is restricted to finite sets.
Here, the alternation-free constraint means that only trivial interactions 
between least- and greatest fixpoint operators are permitted (more 
precise definitions will be provided in Section~\ref{s:mso-mu}).

On the basis of the results by Janin \& Walukiewicz and by Arnold \& 
Niwi{\'n}ski it is natural to conjecture that 
\begin{equation}
\label{eq:weakfinite}
\AFMC \equiv \wmso/{\bis}  \qquad  \text{ (over the class of finitely 
branching trees)},
\end{equation}
where a tree is said to be \emph{finitely branching} if each node has only 
finitely many immediate successors, and the logic $\AFMC$ is the 
alternation-free fragment of $\muML$, or, as we shall explain futher on, the 
`noetherian' version of the modal $\mu$-calculus.
Note that this logic $\AFMC$ is a very natural fragment of the full 
$\mu$-calculus; in particular, since the alternation depth of a fixpoint formula 
is one of the key parameters that determine the complexity of model checking 
algorithms for $\muML$~\cite{calu:deci17},
the alternation-free fragment is of special interest for applications.
Similarly, \emph{weak} $\mso$, the logic featuring on the right-hand side of 
\eqref{eq:weakfinite}, is a well-known variation of standard $\mso$, and it also
has been studied in the setting of applications in software verification (see 
e.g. \cite[Ch. 3]{ALG02}).

In other words, \eqref{eq:weakfinite} is an interesting expressive completeness 
statement, linking two well-known logical systems.
Nevertheless, while we will show that \eqref{eq:weakfinite} holds indeed, our
investigations take a wider scope.
Note that \eqref{eq:weakfinite} only offers a comparison of the logics $\AFMC$
and $\wmso$ on finitely branching trees, whereas \eqref{eq:vB} and \eqref{eq:JW} 
work at the level of arbitrary models. 
In fact, it turns out that the picture in the more general setting is far more
subtle. 
First of all, we know that already at the level of arbitrary trees, the equation
$\AFMC \equiv \wmso/{\bis}$ is \emph{false}, since the class of conversely
well-founded trees, definable by the formula $\mu x.\square x$ of $\AFMC$, is 
not $\wmso$-definable. 
Moreover, whereas $\wmso$ is a fragment of $\smso$ on finitely branching trees,
as soon as we allow for infinite branching the two logics turn out to have
\emph{incomparable} expressive power --- we will discuss the details 
in Section~\ref{s:mso-mu}.

One of the main questions of this work, then, is to clarify the relation between
$\wmso/{\bis}$ and $\AFMC$ on arbitrary LTSs. 
We shall prove that, in this more general setting, the equation 
\eqref{eq:weakfinite} ``splits'' into the following two results, which refer to 
a relatively unknown fragment $\mucML$ of the modal $\mu$-calculus, and 
introduce a new second-order logic $\nmso$.

\begin{theorem}~
   \label{t:11}
\begin{eqnarray}
   \mucML \equiv \wmso/{\bis} \qquad  \text{ (over the class of all LTSs)}. 
\label{eq:mucML=wmso} 
\\ \AFMC \equiv \nmso/{\bis}  \qquad \text{ (over the class of all LTSs)}. 
\label{eq:afmc=nmso}
\end{eqnarray}
\end{theorem}
For the first result \eqref{eq:mucML=wmso}, our strategy is to start from 
$\wmso$ and seek a suitable modal fixpoint logic characterising its 
bisimulation-invariant fragment. 
Second-order quantification $\exists p.\phi$ in $\wmso$ requires $p$ to be 
interpreted over a finite subset of an LTS. 
We identify a notion of \emph{continuity} as the modal counterpart of this 
constraint, and call the resulting logic $\mucML$, the \emph{continuous} 
$\mu$-calculus. 
This fragment of $\muML$, which was introduced in~\cite{bent06:moda}
under the name of ``$\omega$-$\mu$-calculus'', can defined by the same grammar as 
the full $\muML$,
\begin{equation*}
%\label{eq:syntmu}
\phi\ ::= q \mid \neg\phi \mid 
    \phi \lor \phi \mid  \Diamond \phi \mid
    \mu p.\phi' 
\end{equation*}
with the difference that $\phi'$ does not just need to be positive in $p$, but 
also continuous in $p$.
This terminology refers to the fact that $\phi'$ is interpreted by a function 
that is continuous with respect to the Scott topology; as we shall see
in Section~\ref{s:mso-mu}, $p$-continuity can be given a \emph{syntactic} 
characterisation, as a certain fragment of $\muML$ which will be used to define
the logic $\mucML$.

For our second result \eqref{eq:afmc=nmso}, we move in the opposite direction.
That is, we look for a natural second-order logic of which $\AFMC$ is the 
bisimulation-invariant fragment. 
Symmetrically to the case \eqref{eq:mucML=wmso} of $\wmso$ and continuity, a 
crucial aspect is to identify which constraint on second-order quantification
corresponds to the constraint on fixpoint alternation expressed by $\AFMC$. 
Our analysis stems from the observation that, when a formula $\mu p.\phi$ of
$\AFMC$ is satisfied in a tree model $\bbT$, the interpretation of $p$ must be
a subset of a \emph{conversely well-founded} subtree of $\bbT$, because 
alternation-freedom prevents $p$ from occurring in a $\nu$-subformula of $\phi$. 
We introduce the concept of a \emph{noetherian} subset as a generalisation of 
this property from trees to arbitrary LTSs: intuitively, a subset of a LTS 
$\bbS$ is called noetherian if it is a subset of a bundle of paths that does
not contain any infinite ascending chain. 
(Precise definitions will be supplied in Section~\ref{s:mso-mu}.)
The logic $\nmso$ appearing in \eqref{eq:afmc=nmso}, which we call 
\emph{noetherian} second-order logic, is the variation of $\mso$ restricting 
second-order quantification to noetherian subsets.

A unifying perspective over these results can be given through the lens of
K\"onig's lemma, saying that a subset of a tree $\bbT$ is finite precisely 
when it is included in a subtree of $\bbT$ which is both finitely branching 
and conversely well-founded.
In other words, finiteness on trees has two components, a \emph{horizontal}
(finite branching) and a \emph{vertical} (well-foundedness) dimension. 
The bound imposed by $\nmso$-quantification acts only on the \emph{vertical}
dimension, whereas $\wmso$-quantification acts on both. 
It then comes at no surprise that \eqref{eq:mucML=wmso}-\eqref{eq:afmc=nmso}
collapse to \eqref{eq:weakfinite} on finitely branching trees.
The restriction to (unbounded) finitely branching models nullifies the 
difference between noetherian and finite, equating $\wmso$ and $\nmso$ (and 
thus also $\AFMC$ and $\mucML$).

Another interesting observation concerns the relative expressive power of 
$\wmso$ with respect to standard $\mso$. 
As mentioned above, $\wmso$ is \emph{not} strictly weaker than $\smso$ on 
arbitrary LTSs. 
Nonetheless, putting together \eqref{eq:JW} and \eqref{eq:mucML=wmso} reveals 
that $\wmso$ collapses within the boundaries of $\smso$-expressiveness when it 
comes to bisimulation-invariant formulas, because $\mucML$ is strictly weaker
than $\muML$. 
In fact, modulo bisimilarity, $\wmso$ turns out to be even weaker than $\nmso$, 
as $\mucML$ is also a fragment of $\AFMC$. 
In a sense, this new landscape of results tells us that the feature 
distinguishing $\wmso$ from $\smso$/$\nmso$, \emph{viz.} the ability of 
expressing cardinality properties of the horizontal dimension of models, 
disappears once we focus on the bisimulation-invariant part, and thus is not
computationally relevant.

\subsection{Automata-theoretic characterisations}

Janin \& Walukiewicz's proof of \eqref{eq:JW} passes through a characterisation 
of the two logics involved in terms of \emph{parity automata}.
In a nutshell, a parity automaton $\bbA = \tup{A,\tmap,\pmap,a_I}$ processes
LTSs as inputs, according to a transition function $\tmap$ defined in terms of
a so-called \emph{one-step logic} $\oslang(A)$, where the states $A$ of 
$\bbA$ may occur as unary predicates. 
The map $\pmap \colon A \to \mathbb{N}$ assigns to each state a \emph{priority};
if the least priority value occurring infinitely often during the computation is 
even, the input is accepted.
Both $\smso$ and $\muML$ are characterised by classes of parity automata: what 
changes is just the one-step logic, which is, respectively, first-order logic 
with ($\ofoe$) and without ($\ofo$) equality. 
\begin{eqnarray}
\smso & \equiv & \Aut(\ofoe)
 \qquad \text{ (over the class of all trees)}, \label{eq:AutCharMSO}
\\ \muML & \equiv & \Aut(\ofo)
  \qquad \text{ (over the class of all LTSs)}. \label{eq:AutCharMuML}
\end{eqnarray}
This kind of automata-theoretic characterisation, which we believe is of independent interest, also underpins our two correspondence results. As the second main contribution of this paper, we
introduce new classes of parity automata that exactly capture the expressive
power of the second-order languages $\wmso$ and $\nmso$ (over tree models), 
and of the modal languages $\mucML$ and $\mudML$ (over arbitrary models).

Let us start from the simpler case, that is $\nmso$ and $\mudML$.
As mentioned above, the leading intuition for these logics is that they are
constrained in what can be expressed about the \emph{vertical} dimension of 
models. 
In automata-theoretic terms, we translate this constraint into the requirement 
that runs of an automaton can see at most one parity infinitely often: this 
yields the class of so-called \emph{weak} parity 
automata \cite{MullerSaoudiSchupp92}, which we write $\AutW(\oslang)$ for a
given one-step logic $\oslang$. \footnote{%
    Interestingly, \cite{MullerSaoudiSchupp92} introduces the class
    $\AutW(\ofoe)$ in order to shows that it characterises $\wmso$ on 
    finitely branching trees, whence the name of \emph{weak} automata. 
    As discussed above, this correspondence is an ``optical illusion'', due to
    the restricted class of models that are considered, on which $\nmso = 
    \wmso$.
    } 
We shall show:
\begin{theorem}
\begin{eqnarray}
\nmso & \equiv & \AutW(\ofoe)
  \qquad \text{ (over the class of all trees)}, \label{eq:AutCharNmso}
\\ \mudML & \equiv & \AutW(\ofo)\label{eq:AutCharMudML}
  \qquad \text{ (over the class of all LTSs)}.
\end{eqnarray}
\end{theorem}
It is worth to zoom in on our main point of departure from Janin \& Walukiewicz' 
proofs of \eqref{eq:AutCharMSO}-\eqref{eq:AutCharMuML}. 
In the characterisation \eqref{eq:AutCharMSO}, due to \cite{Walukiewicz96}, 
a key step is to show that each automaton in $\Aut(\ofoe)$ can be simulated by
an equivalent \emph{non-deterministic} automaton of the same class. 
This is instrumental in the projection construction, allowing to build an
automaton equivalent to $\exists p.\phi \in \MSO$ starting from an automaton 
for $\phi$. 
Our counterpart \eqref{eq:AutCharNmso} is also based on a simulation theorem. 
However, we cannot proceed in the same manner, as the class $\AutW(\ofoe)$,
unlike $\Aut(\ofoe)$, is \emph{not} closed under non-deterministic simulation.
Thus we devise a different construction, which starting from a weak automaton 
$\bbA$ creates an equivalent automaton $\bbA'$ which acts non-deterministically 
only on a \emph{conversely well-founded} part of each accepted tree.
It turns out that the class $\AutW(\ofoe)$ is closed under this variation of 
the simulation theorem; moreover, the property of $\bbA'$ is precisely what is
needed to make a projection construction that mirrors $\nmso$-quantification.
\medskip

We now consider the automata-theoretic characterisation of $\wmso$ and $\mucML$.
Whereas in \eqref{eq:AutCharNmso}-\eqref{eq:AutCharMudML} the focus was on the 
vertical dimension of a given model, the constraint that we now need to 
translate into automata-theoretic terms concerns both \emph{vertical} and 
\emph{horizontal} dimension. 
Our revision of \eqref{eq:AutCharMSO}-\eqref{eq:AutCharMuML} thus moves on two
different axes. 
The constraint on the vertical dimension is handled analogously to the cases 
\eqref{eq:AutCharNmso}-\eqref{eq:AutCharMudML}, by switching from standard to
\emph{weak} parity automata. 
The constraint on the horizontal dimension requires more work. 
The first problem lies in finding the right one-step logic, which should be able
to express cardinality properties, as $\wmso$ is able to do. 
An obvious candidate would be weak monadic second-order logic itself, or more
precisely, its variation $\owmso$ over the signature of unary predicates 
(corresponding to the automata states).
A very helpful observation from~\cite{vaananen77} is that we can actually work
with an equivalent formalism which is better tailored to our aims.
Indeed, $\owmso \equiv \ofoei$, where $\ofoei$ is the extension of $\ofoe$ with
the generalised quantifier $\qu$, with $\qu x. \phi$ stating the existence of 
\emph{infinitely} many objects satisfying $\phi$. 

At this stage, our candidate automata class for $\wmso$ could be $\AutW(\ofoei)$. However, this fails because $\ofoei$ bears too much expressive power: since it extends $\ofoe$,
we would find that, over tree models, $\AutW(\ofoei)$ extends
$\AutW(\ofoe)$, whereas we already saw that $\AutW(\ofoe) \equiv
\nmso$ is incomparable to $\wmso$.
%
It is here that we crucially involve the notion of \emph{continuity}. 
For a class $\AutW(\oslang)$ of weak parity automata, we call 
\emph{continuous-weak} parity automata, forming a class $\AutWC(\oslang)$, those
satisfying the following additional constraint:
\begin{itemize}
\item 
for every state $a$ with even priority $\pmap(a)$, every one-step formula $\phi
\in \oslang(A)$ defining the transitions from $a$ has to be continuous in all
states $a'$ lying in a cycle with $a$;
dually, if $\pmap(a)$ is odd, every such $\phi$ has to be
$a'$-cocontinuous.\footnote{%
   It is important to stress that, even though continuity is a semantic 
   condition, we have a \emph{syntactic} characterisation of $\ofoei$-formulas
   satisfying it (see \cite{carr:mode18}), meaning that 
   $\AutWC(\ofoei)$ is definable independently of the structures taken as
   input.
   }
\end{itemize}
We can now formulate our characterisation result as follows.
\begin{theorem}
\begin{eqnarray}
\wmso & \equiv & \AutWC(\ofoei)
 \qquad  \text{ (over the class of trees)}, \label{eq:AutCharWmso}
\\ \mucML & \equiv & \AutWC(\ofo)\label{eq:AutCharMucML}
  \qquad \text{ (over the class of all LTSs)}.
\end{eqnarray}
\end{theorem}
Thus automata for $\wmso$ deviate from $\smso$-automata $\Aut(\ofoe)$ on two
different levels: at the global level of the automaton run, because of the 
weakness and continuity constraint, and at the level of the one-step logic 
defining a single transition step. 
Another interesting point stems from pairing
\eqref{eq:AutCharWmso}-\eqref{eq:AutCharMucML} with the expressive completeness
result \eqref{eq:mucML=wmso}: although automata for $\wmso$ are based on a more
powerful one-step logic ($\ofoei$) than those for $\mucML$ ($\ofo$), modulo 
bisimilarity they characterise the same expressiveness.
This connects back to our previous observation, that the ability of $\wmso$ to 
express cardinality properties on the horizontal dimension vanishes in a 
bisimulation-invariant context. 


\subsection{Outline}

It is useful to conclude this introduction with a roadmap of how the various
results are achieved. 
In a nutshell, the two expressive completeness theorems \eqref{eq:mucML=wmso} 
and \eqref{eq:afmc=nmso} will be based respectively on the following two chains 
of equivalences:
\begin{eqnarray}
\AFMC \equiv \mu_{N}\ofo \equiv \AutW(\ofo) \equiv \AutW(\ofoe)/{\bis} \equiv \nmso/{\bis}  \ \text{(over  LTSs)}.\label{eq:chain-afmc=nmso}
\\
	\mucML \equiv \mu_{C}\ofo \equiv \AutWC(\ofo) \equiv \AutWC(\ofoei)/{\bis} \equiv \wmso/{\bis}  \ \text{  (over LTSs)}. \label{eq:chain-mucML=wmso}
\end{eqnarray}

After giving a precise definition of the necessary preliminaries in 
the Sections \ref{sec:prel} and \ref{s:mso-mu}, we proceed as follows. 
First, Section \ref{sec:parityaut} introduces parity automata parametrised over 
a one-step language $\oslang$, both in the standard ($\Aut(\oslang)$), weak 
($\AutW(\oslang)$) and continuous-weak ($\AutWC(\oslang)$) form. 
With Theorems \ref{t:autofor} and \ref{t:fortoaut}, we show that
\begin{equation}\label{sec:outlineFix=Aut}
\mu_{N}\oslang\equiv \AutW(\oslang)  \qquad \qquad \mu_{C}\oslang \equiv \AutWC(\oslang)    \qquad\text{ (over LTSs)}
\end{equation}
where $\mu_{N}\oslang$ and $\mu_{C}\oslang$ are extensions of $\oslang$ with fixpoint operators subject to a ``noetherianess'' and a ``continuity'' constraint respectively. Instantiating \eqref{sec:outlineFix=Aut} yield the second equivalence both in \eqref{eq:chain-afmc=nmso} and \eqref{eq:chain-mucML=wmso}:
\begin{equation*}
	 \mu_{N}\ofo \equiv \AutW(\ofo) \qquad \qquad \mu_{C}\ofo \equiv \AutWC(\ofo)  \qquad \text{ (over LTSs)}.
\end{equation*}
Next, in Section \ref{sec:autwmso}, Theorem~\ref{t:wmsoauto}, we show how to construct from a $\wmso$-formula an equivalent automaton of the class $\AutWC(\ofoei)$. In Section \ref{sec:autnmso}, Theorem~\ref{t:nmsoauto}, we show the analogous characterisation for $\nmso$ and $\AutW(\ofoe)$. These two sections yield part of the last equivalence in \eqref{eq:chain-mucML=wmso} and in \eqref{eq:chain-afmc=nmso} respectively. 
\begin{equation} \label{eq:outlineForToAut}
 \AutW(\ofoe) \geq \nmso \qquad \qquad {\AutWC(\ofoei)} \geq \wmso \qquad \text{ (over trees)}.
 \end{equation}
Notice that, differently from all the other proof pieces, \eqref{eq:outlineForToAut} only holds on trees, because the projection construction for automata relies on the input LTSs being tree shaped. 

Section \ref{sec:fixpointToSO} yields the remaining bit of the automata characterisations. Theorem~\ref{t:mfl2mso} shows
\[ \mu_{N}\ofoe \leq \nmso \qquad \qquad \mu_{C}\ofoei \leq \wmso  \qquad \text{ (over LTSs)},\]
which, paired with \eqref{sec:outlineFix=Aut}, yields
\begin{equation*}%\label{eq:outlineAutToFor}
	 \AutW(\ofoe) \equiv \mu_{N}\ofoe \leq \nmso \quad \AutWC(\ofoei) \equiv \mu_{C}\ofoei \leq \wmso \ \text{ (over LTSs)}.
\end{equation*}
 Putting the last equation and \eqref{eq:outlineForToAut} together we have our automata characterisations
 \begin{equation*}
 \AutW(\ofoe) \equiv \nmso \qquad \qquad  {\AutWC(\ofoei)} \equiv \wmso \qquad \text{ (over trees)}.
 \end{equation*}
 which also yields the rightmost equivalence in \eqref{eq:chain-mucML=wmso} and in \eqref{eq:chain-afmc=nmso}, because any LTS is bisimilar to its tree unraveling.
  \begin{equation*}
 \AutW(\ofoe)/{\bis} \equiv \nmso/{\bis} \qquad \qquad {\AutWC(\ofoei)}/{\bis} \equiv \wmso/{\bis} \qquad \text{ (over LTSs)}.
 \end{equation*}
At last, Section \ref{sec:expresso} is split into two parts. 
First, Theorem \ref{t:mlaut} extends the results in Section \ref{sec:parityaut} 
to complete the following chains of equivalences, yielding the first block 
in \eqref{eq:chain-afmc=nmso} and in \eqref{eq:chain-mucML=wmso}.
\[ 
\AFMC \equiv \mu_{N}\ofo \equiv \AutW(\ofo) \qquad \qquad 
\mucML \equiv \mu_{C}\ofo \equiv \AutWC(\ofo)  \qquad \text{ (over LTSs)}. 
\]
As a final step, Subsection \ref{ss:bisinv} fills the last gap 
in \eqref{eq:chain-afmc=nmso}-\eqref{eq:chain-mucML=wmso} by showing
\[ 
\AutW(\ofo) \equiv \AutW(\ofoe)/{\bis} \qquad \qquad 
\AutWC(\ofo) \equiv \AutWC(\ofoei)/{\bis} \qquad \text{ (over LTSs)}.
\]

  
\subsection{Conference versions and companion paper}
This journal article is based on two conference papers
\cite{DBLP:conf/lics/FacchiniVZ13,DBLP:conf/csl/CarreiroFVZ14},
which were based in their turn on a Master thesis \cite{Zanasi:Thesis:2012}
and a PhD dissertation \cite{carr:frag2015}.
Each of the two conference papers focussed on a single expressive completeness 
theorem between \eqref{eq:mucML=wmso} and \eqref{eq:afmc=nmso}: presenting both
results in a mostly uniform way has required an extensive overhaul, involving
the development of new pieces of theory, as in particular in the
Sections~\ref{sec:parityaut}, \ref{sec:fixpointToSO} and~\ref{sec:expresso}.
%\ref{sec:parity-to-mc}, \ref{sec:mc-to-parity} and \ref{sec:fixpointToSO}. 
All missing proofs of the conference papers are included and the simulation
theorem for $\nmso$- and $\wmso$-automata is simplified, as it is now based
on macro-states that are sets instead of relations. 
Moreover, we amended two technical issues with the characterisation 
$\AFMC \equiv \nmso /{\bis}$ presented in \cite{DBLP:conf/lics/FacchiniVZ13}.
First, the definition of noetherian subset in $\nmso$ has been made more 
precise, in order to prevent potential misunderstandings arising with the
formulation in \cite{DBLP:conf/lics/FacchiniVZ13}. 
Second, as stated in \cite{DBLP:conf/lics/FacchiniVZ13} the expressive 
completeness result was only valid on trees. 
In this version, we extend it to arbitrary LTSs, thanks to the new material 
in Section \ref{sec:fixpointToSO}. 

Finally, our approach depends on model-theoretic results on the three main
one-step logics featuring in this paper: $\ofo$, $\ofoe$ and $\ofoei$.
We believe these results to be of independent interest, and in order to save
some space here, we decided to restrict our discussion of the model theory of
these monadic predicate logics in this paper to a summary.
Full details can be found in the companion paper \cite{carr:mode18}.
