\section{Fixpoint operators and second-order quantifiers}
\label{sec:fixpointToSO}

In this section we will show how to translate some of the mu-calculi that we
encountered until now into the appropriate second-order logics.
Given the equivalence between automata and fixpoint logics that we established 
in Section~\ref{sec:parityaut}, and the embeddings of $\wmso$ and $\nmso$ into,
respectively, the automata classes $\AutWC(\ofoei)$ and $\AutW(\ofoe)$ that
we provided in the Sections~\ref{sec:autwmso} and~\ref{sec:autnmso} for the 
class of tree models, the results here provide the missing link in the 
automata-theoretic characterizations of the monadic second order logics
$\wmso$ and $\nmso$:
\begin{eqnarray*}
   \mu_{C}(\ofoei) \equiv \wmso 
   && \qquad  \text{ (over the class of all tree models)} 
\\ \mu_{N}(\ofoe)  \equiv \nmso 
   && \qquad  \text{ (over the class of all tree models)}. 
\end{eqnarray*}

\subsection{Translating $\mu$-calculi into second-order logics}

More specifically, our aim in this Section is to prove the following result.

\begin{theorem}
\label{t:mfl2mso}
\begin{enumerate}[(1)]
\item
There is an effective translation $(\cdot)^{*}: \mu_{N}\ofoe \to \nmso$
such that $\phi \equiv \phi^{*}$ for every $\phi \in \mu_{N}\ofoe$; that is:
\[
\mu_{N}\ofoe \leq \nmso.
\]
% and as a corollary of this we obtain that 
% \[
% \mu_{N}\ML \leq \nmso.
% \]

\item
There is an effective translation $(\cdot)^{*}: \mu_{C}\ofoei \to \wmso$
such that $\phi \equiv \phi^{*}$ for every $\phi \in \mu_{C}\ofoei$; that is:
\[
\mu_{C}\ofoei \leq \wmso.
\]
\end{enumerate}
\end{theorem}

Two immediate observations on this Theorem are in order.
First, note that we use the same notation $(\cdot)^{*}$ for both translations; 
this should not cause any confusion since the maps agree on formulas belonging 
to their common domain.
Consequently, in the remainder we will speak of a single translation 
$(\cdot)^{*}$.
Second, as the target language of the translation $(\cdot)^{*}$ we will take 
the \emph{two-sorted} version of second-order logic, as discussed in 
section~\ref{sec:prel-so}, and thus we will need Proposition~\ref{p:msovs2mso}
to obtain the result as formulated in Theorem~\ref{t:mfl2mso}, that is,
for the one-sorted versions of \mso.
% although in the Preliminaries we state that in this paper we will work with 
% one-sorted variants of second-order logics.
We reserve a fixed individual variable $v$ for this target language, i.e., 
every formula of the form $\phi^{*}$ will have this $v$ as its unique free 
variable; the equivalence $\phi \equiv \phi^{*}$ is to be understood accordingly.

The translation $(\cdot)^{*}$ will be defined by a straightforward induction on
the complexity of fixpoint formulas.
The two clauses of this definition that deserve some special attention are the
ones related to the fixpoint operators and the modalities.

\paragraph{Fixpoint operators} 
It is important to realise that our clause for the fixpoint operators
\emph{differs} from the one used in the standard inductive translation 
$(\cdot)^{s}$ of $\muML$ into standard $\mso$, where we would inductively
translate $(\mu p. \phi)^{*}$ as
\begin{equation}
\label{eq:st}
% \exists p\, \big(   p(x) \land \forall y\, (p(y) \to \phi^{s}_{y}) \big)
\forall p\, \big( \forall w\, (\phi^{*}[w/v] \to p(w)) \to p(v) \big),
\end{equation}
which states that $v$ belongs to any prefixpoint of $\phi$ with respect to $p$.
To understand the problem with this translation in the current context, suppose,
for instance, that we want to translate some continuous $\mu$-calculus into
$\wmso$.
Then the formula in \eqref{eq:st} expresses that $v$ 
belongs to the intersection of all \emph{finite} prefixpoints of $\phi$, whereas
the least fixpoint is identical to the intersection of \emph{all} prefixpoints.
As a result, \eqref{eq:st} does not give the right translation for the formula 
$\mu p.\phi$ into \wmso.

To overcome this problem, we will prove that least fixpoints in restricted 
calculi like $\mu_{N}\ofoe$, $\mu_{C}\ofoei$ and many others, in fact satisfy a
rather special property, which enables an alternative translation.
We need the following definition to formulate this property.

\begin{definition}
\label{d:rst}
Let $F: \pow(S)\to \pow(S)$ be a functional; for a given $X \subseteq S$ we define
the \emph{restricted} map $F_{\rst{X}}: \pow(S)\to \pow(S)$ by putting 
$F_{\rst{X}}(Y) \isdef FY \cap X$.
\textcolor{red}{
In case $F$ is monotone, we will use $\LFP.F$ to denote its least fixpoint.
} 
\end{definition}


The observations formulated in the proposition below provide the crucial insight
underlying our embedding of various alternation-free and continuous 
$\mu$-calculi into, respectively, $\nmso$ and $\wmso$. 

\begin{proposition}
\label{p:afmc-rstGen}
\label{p:keyfix}
Let $\bbS$ be an LTS, and let $r$ be a point in $\bbS$.
\begin{enumerate}[(1)]
\item
 For any formula $\phi$ with $\mu p. \phi \in \mu_{N}\ofoe$ we have
\begin{equation}
\label{eq:foe-d}
r \in \ext{\mu p.\phi}^{\bbS} \text{ iff there is a noetherian set $X$ such 
that } r \in \LFP. (\phi^{\bbS}_{p})_{\rst{X}}.
\end{equation}

\item
For any formula $\phi$ with $\mu p. \phi \in \mu_{C}\ofoei$ we have
\begin{equation}
\label{eq:foei-c}
r \in \ext{\mu p.\phi}^{\bbS} \text{ iff there is a finite set $X$ such 
that } r \in \LFP. (\phi^{\bbS}_{p})_{\rst{X}}.
\end{equation}
\end{enumerate}
\end{proposition}

\begin{remark}
In fact, the statements in Proposition~\ref{p:keyfix} can be generalised to the
setting of a fixpoint logic $\mu\oslang$ associated with an arbitrary one-step 
language $\oslang$.
\end{remark}

% \textcolor{blue}{For the right-to-left direction of both \eqref{eq:foe-d} and 
% \eqref{eq:foei-c} we reason as follows. 
% Let $F,G:  \pow(S)\to \pow(S)$ be monotone, and such that 
% $F(X) \subseteq G(X)$, for every $X \in  \pow(S)$. Then 
% any prefixpoint of $G$ is also a prefixpoint of $F$. In particular, this means that $\LFP. F_{\rst{X}}\subseteq 
% \LFP.F$, for every subset $X \subseteq S$.}

\textcolor{red}{
The right-to-left direction of both \eqref{eq:foe-d} and \eqref{eq:foei-c} 
follow from the following, more general, statement, which can be proved by a 
routine argument.
%
\begin{proposition}
% \label{p:afmc-1}
\label{p:rstfix}
Let $F,G:  \pow(S)\to \pow(S)$ be monotone, and assume that 
$F(Y) \subseteq G(Y)$, for every $Y \in  \pow(S)$.
Then $\LFP. F \subseteq \LFP.G$.
\end{proposition}
}

%The right-to-left direction of both \eqref{eq:foe-d} and \eqref{eq:foei-c} follow
%from the following, more general, statement.%, which can be proved by a routine  transfinite induction argument.
%
%\begin{proposition}
%% \label{p:afmc-1}
%\label{p:rstfix}
%Let $F:  \pow(S)\to \pow(S)$ be monotone.
%Then for every subset $X \subseteq S$ it holds that $\LFP. F_{\rst{X}}\subseteq 
%\LFP.F$.
%\end{proposition}


The left-to-right direction of \eqref{eq:foe-d} and \eqref{eq:foei-c} 
will be proved in the next two subsections.
Note that in the continuous case we will in fact prove a slightly stronger
result, which applies to \emph{arbitrary} continuous functionals.
% , we discuss this case separately.

\newcommand{\PRE}{\mathit{PRE}}
The point of Proposition~\ref{p:keyfix} is that it naturally suggests the
following translation for the least fixpoint operator, as a subtle but important 
variation of \eqref{eq:st}:

\begin{equation}
\label{eq:trlmu}
(\mu p. \varphi)^{*} \df 
   \exists q\,\Big(\forall  p \sse q.\,
      \big(p \in \PRE((\varphi^{\bbS}_{p})_{\rst{q}}) \to p(v)\big)\Big),
\end{equation}
where $p \in \PRE((\varphi^{\bbS}_{p})_{\rst{q}})$ expresses that $p \sse q$ 
is a prefixpoint of the map $(\phi^{\bbS}_{p})_{\rst{q}}$, that is:
\[
p  \in \PRE((\varphi^{\bbS}_{p})_{\rst{q}}) \df
\forall w\, \Big(
( q(w) \land \varphi^{*}[w/v]) \to p(w)
\Big).
\]

\paragraph{Modalities}
Finally, before we can give the definition of the translation $(\cdot)^{*}$, we
briefly discuss the clause involving the modalities.
Here we need to understand the role of the \emph{one-step formulas} in the 
translation.
First an auxiliary definition.

\begin{definition}\label{def:exp}
Let $\bbS = \tup{T,R,\tscolors, s_I}$ be a $\pprop$-LTS, let $A$ be a set of
new variables, and let $V: A \to \pow(X)$ be a valuation on a subset $X\subseteq
T$. 
The $(\pprop\cup A)$-LTS $\bbS^{V}\isdef \tup{T,R,\tscolors^V, s_I}$, given by 
defining the marking $\tscolors^V: T \to \pow{(\pprop \cup A)}$ where
\[\tscolors^V(s)\isdef  
\begin{cases} \tscolors(s) & \text{ if } s \notin X \\
\tscolors(s) \cup \{ a \in A \mid s \in V(a)\}& \text{ else,}
\end{cases}\]
is called the \emph{$V$-expansion} of $\bbS$.
\end{definition}

The following proposition states that at the one-step level, the formulas that 
provide the semantics of the modalities of $\mu\ofoe$ and $\mu\ofoei$ can indeed 
be translated into, respectively $\nmso$ and $\wmso$.

\begin{proposition}
\label{p:1trl}
There is a translation $(\cdot)^{\dag}: \ofoei(A) \to \wmso$ such that for every 
model $\bbS$ and every valuation $V: A \to \pow(R[s_{I}])$:
\[
(R[s_{I}],V) \models \al \text{ iff } \bbS^{V} \models \al^{\dag}[s_{I}].
\]
Moreover, $(\cdot)^{\dag}$ restricts to first-order logic, i.e., $\al^{\dag}$ is
a first-order formula if $\al \in \ofoe$.
\end{proposition}

% \btbs
% \item
% Somehow clarify in the notation that $\al^{\dag}$ has exactly one free 
% individual variable, $v$.
% \etbs

\begin{proof}
Basically, the translation $(\cdot)^{\dag}$ \emph{restricts} all quantifiers
to the collection of successors of $v$.
In other words, $(\cdot)^{\dag}$ is the identity on basic formulas, it commutes
with the propositional connectives, and for the quantifiers $\exists$ and $\qu$
we define:
\[\begin{array}{lll}
(\exists x\, \al)^{\dag} &\df& \exists x\, (Rvx \land \al^{\dag})
\\ (\qu x\, \al)^{\dag}  &\df& \forall p \exists x\, (Rvx \land \neg p(x) 
    \land \al^{\dag})
\end{array}\]
We leave it for the reader to verify the correctness of this definition ---
observe that the clause for the infinity quantifier $\qu$ is based on the 
equivalence between $\wmso$ and $\foei$, established by 
V\"a\"an\"anen~\cite{vaananen77}.
\end{proof}

\noindent
We are now ready to define the translation used in the main result of this 
section.

\begin{definition}
By an induction on the complexity of formulas we define the following 
translation $(\cdot)^{*}$ from $\mu\foei$-formulas to formulas of monadic 
second-order logic:
\[\begin{array}{lll}
   p^{*} &\df& p(v)
\\ (\neg\phi)^{*}        &\df& \neg \phi^{*}
\\ (\phi\lor\psi)^{*}    &\df& \phi^{*} \lor \psi^{*}
\\ (\nxt{\al}(\ol{\phi}))^{*} &\df& \al^{\dag}[\phi_{i}^{*}/a_{i} \mid i \in I],
\end{array}\]
where $\al^{\dag}$ is as in Proposition~\ref{p:1trl}, and $[\phi_{i}^{*}/a_{i}
\mid i \in I]$ is the substitution that replaces every occurrence of an atomic
formula of the form $a_{i}(x)$ with the formula $\phi_{i}^{*}(x)$ (i.e. the 
formula $\phi_{i}^{*}$, but with the free variable $v$ substituted by $x$).

Finally, the inductive clause for a formula of the form $\mu p.\phi$ is given
as in \eqref{eq:trlmu}.
\end{definition}

\begin{proofof}{Theorem~\ref{t:mfl2mso}}
%\btbs
%\item
%For both parts of the theorem we need to verify that 
%\\ (i) the translation $(\cdot)^{*}$ lands in the correct language, 
%\\ (ii) the translation $(\cdot)^{*}$ is truth preserving and
%\\ (iii) the statement about the modal language indeed follows from the 
%   main statement.
%\etbs
First of all, it is clear that in both cases the translation $(\cdot)^{*}$ lands 
in the correct language.
For both parts of the theorem, we thence prove that $(\cdot)^{*}$ is truth
preserving by a straightforward formula induction.
E.g., for part (2) we need to show that, for an arbitrary formula $\phi\in
\mu_{C}\ofoei$ and an arbitrary model $\bbS$:
\begin{equation}
\label{eq:xxxx1}
\bbS \mmodels \phi \text{ iff } \bbS \models \phi^{*}[s_{I}].
\end{equation}

As discussed in the main text, the two critical cases concern the inductive 
steps for the modalities and the least fixpoint operators. 
Let $ \oslang^{+} \in \{\ofoe,\ofoei\}$.
We start verifying the case of modalities. 
Hence, consider the formula $\nxt{\al}(\varphi_{1},\ldots,\varphi_{n})$ with 
$\al(a_{1},\ldots,a_{n}) \in \oslang^{+}$. 
By induction hypothesis, $\phi_\ell \equiv \phi^{*}_\ell$, for $\ell=1,\dots, n$.
Now, let $\bbS$ be a transition system. We have that
\begin{align*}
\bbS \mmodels \nxt{\al}(\varphi_{1},\ldots,\varphi_{n}) \text{ iff } 
  & (R[s_{I}],V_{\overline{\varphi}}) \models \al(a_{1},\ldots,a_{n})  
  & \text{( by \eqref{eq:mumod})}
\\ \text{ iff } 
  & \bbS^{V_{\overline{\varphi}}} \models \al^{\dag}[s_{I}] 
  & \text{( by Prop. \ref{p:1trl})}
\\ \text{ iff } 
  & \bbS \models \al^{\dag}[\phi_{i}^{*}/a_{i} \mid i \in I][s_{I}] 
  & \text{( by \eqref{eq:valmod}, Def. \ref{def:exp} and IH)}
\end{align*}

The inductive step for the least fixpoint operator will be justified by 
Proposition~\ref{p:keyfix}.
In more detail, given a formula of the form $\mu x. \psi \in\mu_{Y} \oslang^{+}$,
with $Y=D$ for $\oslang^{+} =\ofoe$, and $Y=C$ for $\oslang^{+} =\ofoei$, 
consider the following chain of equivalences:
\begin{align*}
% s_{I} \in \ext{\mu p.\psi}^{\bbS} \text{ iff } 
  & s_{I} \in \ext{\mu p.\psi}^{\bbS} 
\\ \text{ iff } 
  & s_{I} \in \LFP. (\psi^{\bbS}_{p})_{\rst{Q}} \text{ for some }
        \begin{cases} \text{ finite} & \\ \text{ noetherian} & 
	\end{cases} 
    \text{set } Q 
  & \text{( by \eqref{eq:foe-d}/\eqref{eq:foei-c})}
\\ \text{ iff } 
  &  s_{I} \in \bigcap \Big\{ P \subseteq Q \mid P \in 
         \PRE((\psi^{\bbS}_{p})_{\rst{Q}})\Big\} 
     \text{ for some }
        \begin{cases} \text{ finite} & \\ \text{ noetherian} & 
	\end{cases} 
     \text{set } Q 
\\ \text{ iff } & 
    \bbS \models \exists q.\, \Big(\forall p \subseteq q.\,
       \big(p \in \PRE((\psi^{\bbS}_{p})_{\rst{q}}) \to p(s_{I})\big)
       \Big)
\\ \text{ iff } & 
    \bbS \models (\mu p. \psi)^{*}[s_{I}].
   & (\text{IH})
\end{align*}
This concludes the proof of \eqref{eq:xxxx1}.
% \btbs
% \item
% We finally verify that the statement about the modal languages follows from the
% main statement of the Theorem.
% In order to do this, it is enough to check that for any set $Q$ of propositional
% variables, both $\noe{\mu\ML}{Q}\subseteq\noe{\mu\ofoei}{Q}$ and 
% $\conoe{\mu\ML}{Q}\subseteq\conoe{\mu\ofoei}{Q}$, and 
% $\cont{\mu\ML}{Q}\subseteq\cont{\mu\ofoei}{Q}$ and 
% $\cocont{\mu\ML}{Q}\subseteq\cocont{\mu\ofoei}{Q}$ hold. 
% But this follows by a straightforward induction on the structure of a formula
% by noticing that $\Diamond \varphi = \nxt{\al}(\varphi)$ for $\al(a)\isdef 
% \exists x.a(x) \in \ofoei \cap \ofoe$ and $\Box \varphi = \nxt{\al'}(\varphi)$
% for $\al'(a)\isdef \forall x.a(x) \in \ofoei \cap \ofoe$.
% %This conclude the proof of the Theorem.
% \etbs
\end{proofof}
\subsection{Fixpoints of continuous maps}

It is well-known that continuous functionals are \emph{constructive}.
That is, if we construct the least fixpoint of a continuous functional $F: 
\pow(S) \to \pow(S)$ using the ordinal approximation $\nada, F\nada, F^2\nada, 
\ldots, F^{\al}\nada, \ldots$, then we reach convergence after at most $\omega$
many steps, implying that $\LFP. F = F^{\omega}\nada$.
We will see now that this fact can be strengthened to the following observation,
which is the crucial result needed in the proof of Proposition~\ref{p:keyfix}.

\begin{theorem}
\label{t:fixcont}
Let $F: \pow(S)\to \pow(S)$ be a continuous functional.
Then for any $s \in S$:
\begin{equation}
\label{eq:fixcont}
s \in \LFP. F \text{ iff }
s \in \LFP.F_{\rst{X}}, \text{ for some finite } X \subseteq S.
\end{equation}
\end{theorem}

\begin{proof}
The direction from right to left of \eqref{eq:fixcont} is a special case of 
Proposition~\ref{p:rstfix}.
For the opposite direction of \eqref{eq:fixcont} a bit more work is needed.
Assume that $s \in \LFP. F$; we claim that there are sets $U_{1},\ldots,U_{n}$,
for some $n \in \omega$, such that $s \in U_{n}$, $U_{1} \sse_{\omega} F(\nada)$,
and $U_{i+1} \sse_{\omega} F(U_{i})$, for all $i$ with $1 \leq i < n$.

To see this, first observe that since $F$ is continuous, we have $\LFP. F = 
F^{\omega}(\nada) = \bigcup_{n\in\omega}F^{n}(\nada)$, and so we may take $n$ to
be the least natural number such that $s \in F^{n}(\nada)$.
By a downward induction we now define sets $U_{n},\ldots,U_{1}$, with $U_{i} 
\sse F^{i}(\nada)$ for each $i$. 
We set up the induction by putting $U_{n} \isdef \{ s \}$, then $U_{n}
\sse F^{n}(\nada)$ by our assumption on $n$.
For $i<n$, we define $U_{i}$ as follows.
Using the inductive fact that $U_{i+1} \sse_{\omega} F^{i+1}(\nada) = 
F(F^{i}(\nada))$, it follows by continuity of $F$ that for each $u \in U_{i+1}$
there is a set $V_{u} \sse_{\omega} F^{i}(\nada)$ such that $u \in F(V_{u})$.
We then define $U_{i} \isdef \bigcup \{ V_{u} \mid u \in U_{i+1} \}$,
so that clearly $U_{i+1} \sse_{\omega} F(U_{i})$ and $U_{i} \sse_{\omega}
F^{i}(\nada)$.
Continuing like this, ultimately we arrive at stage $i=1$ where we find
$U_{1} \sse F(\nada)$ as required.

Finally, given the sequence $U_{n},\ldots,U_{1}$, we define 
\[
X \isdef  \bigcup_{0<i\leq n} U_{i}.
\]
It is then straightforward to prove that $U_{i} \sse \LFP. F_{\rst{X}}$, for each 
$i$ with $0<i\leq n$, and so in particular we find that $s \in U_{n} \sse \LFP.
F_{\rst{X}}$.
This finishes the proof of the implication from left to right in 
\eqref{eq:fixcont}.
\end{proof}

\noindent
As an almost immediate corollary of this result we obtain the second part of 
Proposition~\ref{p:keyfix}.

\begin{proofof}{Proposition~\ref{p:keyfix}(2)}
Take an arbitrary formula $\mu p. \phi \in \mu_{C}\ofoei$, then by definition 
we have $\phi \in \mu_{C}\ofoei \cap \cont{\mu\ofoei}{p}$.
But it follows from a routine inductive proof that every formula $\psi \in 
\mu_{C}\ofoei \cap \cont{\mu\ofoei}{\qprop}$ is continuous in each variable 
in $\qprop$.
Thus $\phi$ is continuous in $p$, and so the result is immediate by 
Theorem~\ref{t:fixcont}.
\end{proofof}


% !TEX root = ../00CFVZ_TOCL.tex

\subsection{Fixpoints of noetherian maps}

We will now see how to prove Proposition~\ref{p:keyfix}(1), which is the key 
result that we need to embed alternation-free $\mu$-calculi such as 
$\mu_{N}\ofoe$ and $\mu_{N}\ML$ into noetherian second-order logic.
Perhaps suprisingly, this case is slightly more subtle than the characterisation of
fixpoints of continuous maps.

We start with stating some auxiliary definitions and results on monotone 
functionals, starting with a game-theoretic characterisation of their least
fixpoints~\cite{Ven08}.

\begin{definition}
\label{d:unfgame}
Given a monotone functional $F: \pow(S)\to \pow(S)$ we define the 
\emph{unfolding game} $\UG_{F}$ as follows:
\begin{itemize}
\item at any position $s \in S$, $\eloise$ needs to pick a set $X$ such that 
$s \in FX$;
\item at any position $X \in \pow(S)$, $\abelard$ needs to pick an element of 
$X$
\item all infinite matches are won by $\abelard$.
\end{itemize}
A positional strategy $\ystrat: S \to \pow(S)$ for $\eloise$ in $\UG_{F}$ is 
\emph{descending} if, for all ordinals $\alpha$,
\begin{equation}
\label{eq:dec}
s \in F^{\alpha+1}(\nada) \text{ implies } \ystrat(s) \sse F^{\alpha}(\nada).
\end{equation}
\end{definition}

It is not the case that \emph{all} positional winning strategies for $\eloise$ 
in $\UG_{F}$ are descending, but the next result shows that there always is one.

\begin{proposition}
\label{p:unfg}
Let $F: \pow(S)\to \pow(S)$ be a monotone functional. 
\begin{enumerate}[(1)]
\item
For all $s \in S$, $s \in \win_{\eloise}(\UG_{F})$ iff $s \in \LFP. F$;
\item 
If $s \in \LFP. F$, then \eloise has a descending winning strategy in 
$\UG_{F}@s$.
\end{enumerate}
\end{proposition}

\begin{proof}
Point (1) corresponds to \cite[Theorem 3.14(2)]{Ven08}.
For part (2) one can simply take the following strategy.
Given $s \in \LFP.F$, let $\alpha$ be the least ordinal such that $s \in 
F^{\alpha}(\nada)$; it is easy to see that $\alpha$ must be a successor ordinal,
say $\alpha = \beta + 1$. 
Now simply put $\ystrat(s) \isdef  F^{\beta}(\nada)$.
\end{proof}

\begin{definition}
\label{d:str-tree}
Let $F: \pow(S)\to \pow(S)$ be a monotone functional, let $\ystrat$ be a 
positional winning strategy for $\eloise$ in $\UG_{F}$, and let $r \in S$. 
Define $T_{\ystrat,r} \sse S$ to be the set of states in $S$ that are 
$\ystrat$-reachable in $\UG_{F}@r$.
This set has a tree structure induced by the map $\ystrat$ itself, where the 
children of $s \in T_{\ystrat,r}$ are given by the set $\ystrat(s)$; we will
refer to $T_{\ystrat,r}$ as the \emph{strategy tree} of
$\ystrat$.
\end{definition}
Note that a strategy tree $T_{\ystrat,r}$ will have no infinite paths, since we
define the notion only for a \emph{winning} strategy $\ystrat$.

\begin{proposition}
\label{p:afmc-2}
Let $F: \pow(S)\to \pow(S)$ be a monotone functional, let $r \in S$, and let 
$\ystrat$ be a descending winning strategy for $\eloise$ in $\UG_{F}$.
Then
\begin{equation}
\label{eq:afmc3}
r \in \LFP. F \text{ implies } r \in \LFP. F_{\rst{T_{\ystrat,r}}}.
\end{equation}
\end{proposition}

\begin{proof}
Let $F,r$ and $\ystrat$ be as in the formulation of the proposition.
Assume that $r \in \LFP. F$, then clearly $r \in F^{\al}(\nada)$ for some 
ordinal $\al$; furthermore, $T_{\ystrat,r}$ is defined and clearly we have 
$r \in T_{\ystrat,r}$.
Abbreviate $T \isdef T_{\ystrat,r}$.
It then suffices to show that for all ordinals $\alpha$ we have
\begin{equation}
\label{eq:unf1}
F^{\al}(\nada) \cap T \sse (F_{\rst{T}})^{\alpha}(\nada).
\end{equation}
We will prove \eqref{eq:unf1} by transfinite induction.
The base case, where $\alpha = 0$, and the inductive case where $\alpha$ is a 
limit ordinal are straightforward, so we focus on the case where $\alpha$ is a 
successor ordinal, say $\alpha = \beta +1$.
Take an arbitrary state $u \in F^{\beta+1}(\nada) \cap T$, then we 
find $\ystrat(u) \sse F^{\beta}(\nada)$ by our assumption \eqref{eq:dec}, and 
$\ystrat(u) \sse T$ by definition of $T$.
Then the induction hypothesis yields that 
$\ystrat(u) \sse (F_{\rst{T}})^{\beta}(\nada)$, and so we have 
$\ystrat(u) \sse (F_{\rst{T}})^{\beta}(\nada) \cap T$.
But since $\ystrat$ is a winning strategy, and $u$ is a winning position for 
$\eloise$ in $\UG_{F}$ by Claim~\ref{p:unfg}(i), $\ystrat(u)$ is a
legitimate move for $\eloise$, and so we have $u \in F(\ystrat(u))$.
Thus by monotonicity of $F$ we obtain $u \in 
F((F_{\rst{T}})^{\beta}(\nada) \cap T)$, and since $u \in T$ 
by assumption, this means that $u \in (F_{\rst{T}})^{\beta+1}(\nada)$ as 
required.
\end{proof}

We now turn to the specific case where we consider the least fixed point of a 
functional $F$ which is induced by some formula $\phi(p) \in 
\mu_{N}\oslang$ on some LTS $\bbS$.   
By Proposition \ref{p:unfg} and Fact~\ref{f:adeqmu}, $\eloise$ has a winning
strategy in  $\egame(\mu p.\phi(p),\bbS)@(\mu p.\phi(p),s)$ if and only if she
has a winning strategy in $\UG_{F}@s$ too, where $F \isdef  \phi_{p}^{\bbS}$ is the 
monotone functional defined by $\phi(p)$. 
The next Proposition  makes this correspondence explicit when $\oslang =
\foe$. 

%\newpage

First, we need to introduce some auxiliary concepts and notations.
Given  a winning strategy   $\ystrat$ for $\eloise$ in $\egame(\mu p. \phi,
\bbS)@(\mu p. \phi,s)$, we denote by $B(\ystrat)$ the set of all finite 
$\ystrat$-guided, possibly partial, matches in  $\egame(\psi,\bbS)@(\psi,s)$ in
which no position of the form $(\nu q. \psi, r)$ is visited. Let 
$\ystrat$ be a positional winning strategies for $\eloise$ in $\UG_F@s$ and 
$\ystrat'$ a winning strategy  for her in $\egame(\mu p.\phi,\bbS)@(\mu p.\phi,
s)$. We call $\ystrat$ and  $\ystrat'$  \emph{compatible} if
each point in $T_{\ystrat,s}$ occurs on some path belonging to $B(\ystrat')$. 

\begin{proposition}\label{p:unfold=evalgame2}
Let $\phi(p) \in \mu_{N}\foe{p}$ and $s \in \ext{\mu p.\phi}^{\bbS}$. 
Then there is a descending winning strategy for $\eloise$ in $\UG_F@s$ 
compatible with a  winning strategy for $\eloise$ in 
$\egame(\mu p.\phi,\bbS)@(\mu p.\phi,s)$
\end{proposition}

\begin{proof}
Let $F \isdef  \phi_{p}^{\bbS}$ be the monotone functional defined by $\phi(p)$.
From $s \in \ext{\mu p.\phi}^{\bbS}$, we get that $s \in \LFP. F$.  
Applying Proposition~\ref{p:unfg} to the fact that $s \in \LFP. F$ yields that 
$\eloise$ has a descending winning strategy $\ystrat: S \to \pow{(S)}$ in $\UG_{F}@s$. 
We define $\eloise$'s strategies $\ystrat'$  in
$\egame(\mu p.\phi,\bbS)@(\mu p.\phi,s)$, and $\ystrat^*$ in $\UG_{F}@s$ as follows:
\begin{enumerate}
\item 
In the evaluation games $\egame$, after the initial automatic move, the position 
of the match is $(\phi,s)$; there $\eloise$ first plays her positional winning 
strategy $\ystrat_s$ from $\egame(\phi(p),\bbS[p \mapsto \ystrat(s)])@(\phi(p),
s)$, and we define her move $\ystrat^*(s)$ in the unfolding game $\UG$ as the 
set of all nodes $t \in \ystrat(s)$ such that there is a $\ystrat_s$-guided 
match in $B(\ystrat_s)$ whose last position is $(p,t)$.
\item 
Each time a position $(p,t)$ is reached in the evaluation games $\egame$, 
distinguish cases:
\begin{enumerate}
%necessarily 
\item 
if $t \in \win_{\eloise}(\UG_{F})$, then $\eloise$ continues with the positional
winning strategy $\ystrat_t$ from $\egame(\phi(p),\bbS[p \mapsto
\ystrat(t)])@(\phi(p),t)$, and we define her move $\ystrat^*(t)$ in $\UG$ as the 
set of all nodes $w \in \ystrat(t)$ such that there is a $\ystrat_t$-guided
match in $B(\ystrat_t)$ whose last position $(p,w)$;
\item 
if $t \notin \win_{\eloise}(\UG_{F})$, then $\eloise$ continues with a random 
positional strategy and we define $\ystrat^*(t)\isdef \nada$.
\end{enumerate}
\item 
For any position $(p,t)$ that was not reached in the previous steps, $\eloise$ 
sets $\ystrat^*(t)\isdef \nada$.
\end{enumerate}
By construction, $\ystrat'$ and  $\ystrat^*$ are compatible. 
Moreover, $\ystrat^*(t) \subseteq \ystrat(t)$, for $t \in S$, meaning that 
$\ystrat^*$ is descending.
We verify that both $\ystrat'$ and  $\ystrat^*$ are actually winning strategies
for $\eloise$ in the respective games.

First of all, observe that every position of the form $(p,t)$ reached during
a $\ystrat'$-guided match, we have $t \in \win_{\eloise}(\UG_{F})$. 
This can be proved by induction on the number of position of the form $(p,t)$
visited during an $\ystrat'$-guided match. 
For the inductive step, assume $w \in \win_{\eloise}(\UG_{F})$. 
Hence $\ystrat_w$ is winning for $\eloise$ in  $\egame(\phi,\bbS[p \mapsto 
  \ystrat(w)])@(\phi,w)$. 
This means that if a position of the form $(p, t)$ is reached, the variable $p$ 
must be true at $t$ in the model $\bbS[p \mapsto \ystrat(w)]$, meaning that it
belongs to the set $\ystrat(w)$.
By assumption $\ystrat$ is a winning strategy for $\eloise$ in $\UG_F$, and 
therefore any element of $\ystrat(w)$ is again a member of the set 
$\win_{\eloise}(\UG_{F})$. 

Finally, let $\pi$ be an arbitrary $\ystrat'$-guided match of $\egame(\phi,
\bbS[p \mapsto \ystrat(w)])@(\phi,w)$. 
We verify that $\pi$ is winning for $\eloise$. 
First observe that since $\ystrat$ is winning for her in $\UG_F@s$, the fixpoint 
variable $p$ is unfolded only finitely many times during $\pi$. 
Let $(p,t)$ be the last basic position in $\pi$ where $p$ occurs. 
Then from now on $\ystrat'$ and $\ystrat_t$ coincide, yielding  that the match is 
winning for $\eloise$. 

%We define \[P_{\last(\pi)}\isdef \{ (p,w) \mid \exists \pi' \in B(\ystrat'), \pi'=\pi \rho(p,w), p\text{ does not occur in }\rho\}.\]
%Recall that $B(\ystrat')$ is the set of all finite $\ystrat'$-guided, possibly partial, matches in 
%$\egame$ in which no position of the form 
%$(\nu q. \psi, r)$ is visited.
%\\
%It is not hard to derive from the positionality of $\ystrat'$ that for every $\pi, 
%\pi' \in B(\ystrat')$ such that $\last(\pi)=\last(\pi')$, it holds that $\pi\rho 
%\in B(\ystrat')$ if and only if $\pi'\rho \in B(\ystrat')$, for every finite match 
%$\rho$.
%\\
%Now consider  the  strategy $\ystrat^*$ in $\UG_{F}@s$ defined by 
%\[
%\ystrat^*(t) \isdef \{ w \mid (p,w) \in P_{\last(\pi)}, \pi\in B(\ystrat'), 
%\last(\pi)=(\mu p. \phi(p), t)\}.
%\]
%\\
%Compatibility being immediate, 
We finally verify that $\ystrat^*$ is winning for $\eloise$ in the unfolding
game $\UG_F@s$. 
First of all, since $\ystrat'$ is winning, $B(\ystrat')$ does not contain an
infinite ascending chain of $\ystrat'$-guided matches, and thence any
$\ystrat^*$-guided match in $\UG_{F}@s$ is finite. 
It therefore remains to verify that for every $\ystrat^*$-guided match $\pi$ 
in $\UG_{F}@s$ such that $\last(\pi)$ is an $\eloise$ position, she can always
move.
We do it by induction on the length of a $\ystrat^*$-guided match. 
At each step, we use compatibility and thus keep track of the corresponding 
position in the evaluation game $\egame(\mu p.\phi,\bbS)@(\mu p.\phi,s)$. 
The initial position for her is $s \in S$. Notice that $\ystrat^*(s) = 
\ystrat(s) \cap B(\xi')$ and therefore $\ystrat'$ corresponds to $\ystrat_s$
on $\egame(\phi(p),\bbS[p \mapsto \ystrat^*(s)])@(\phi(p),s)$ and it is 
therefore winning for $\eloise$.
In particular, this means  that $s \in F(\ystrat^*(s))$.
Hence, as initial move, $\eloise$ is allowed to play $\ystrat^*(s)$. 
Moreover any subsequent choice for $\abelard$ is such that there is a winning
match $\pi \in B(\xi_s)$ for $\eloise$ such that $\last(\pi)=(p,w)$. 
For the induction step, assume $\abelard$ has chosen $ t \in \ystrat^*(w)$, 
where $\ystrat^*(w) = \ystrat(w) \cap B(\xi')$, $\ystrat'$ corresponds to the
winning strategy $\ystrat_w$ on $\egame(\phi(p),\bbS[p \mapsto 
\ystrat^*(w)])@(\phi(p),w)$, and there is a winning match $\pi \in B(\xi_w)$ 
for $\eloise$ such that $\last(\pi)=(p,w)$. 
By construction, $\ystrat'$ corresponds to the winning strategy $\ystrat_t$ 
for $\eloise$ on $\egame(\phi(p),\bbS[p \mapsto \ystrat(t)])@(\phi(p),t)$. 
Because $\ystrat^*(t)= \ystrat(s) \cap B(\xi')$, $\ystrat_t$ is also winning
for her in $\egame(\phi(p),\bbS[p \mapsto \ystrat^*(t)])@(\phi(p),t)$, meaning
that $s \in F(\ystrat^*(s))$. 
The move $\ystrat^*(t)$ is therefore admissible, and any subsequent choice for
$\abelard$ is such that there is a winning match $\pi \in B(\xi_t)$ for 
$\eloise$ with $\last(\pi)=(p,w)$.
\end{proof}

%Notice that the evaluation game for the modal $\mu$ calculus can be easily adapted to $\mu_{N}\foe$ and provide a corresponding Adequacy Theorem (Proposition \ref{p:unfold=evalgame}). In particular, we would have that 
%$\eloise$ has a winning strategy in  $\egame(\mu p.\phi(p),\bbS)@(\mu p.\phi(p),s)$ if and only if $s \in \ext{\mu p.\phi(p)}^{\bbS}$, with $\phi(p) \in \mu_{N}\foe{p}$.

\begin{proofof}{Proposition~\ref{p:keyfix}(1)}
%    
%\btbs
%\item
%the argument below needs to be thoroughly checked.
%\etbs
Let $\bbS$ be an LTS and $\phi(p) \in \mu_{N}\ofoe{p}$. 

The right-to-left direction of \eqref{eq:foe-d} being proved by 
Proposition \ref{p:rstfix}, we check the left-to-right direction.
We first verify that winning strategies in evaluation games for noetherian 
fixpoint formulas naturally induce bundles. 
More precisely:
\medskip
%Given  a winning strategy   $\ystrat$ for $\eloise$ in $\egame(\mu p. \phi,\bbS)@(\mu p. \phi,s)$,  let $B(\ystrat)$ be the set of all finite $\ystrat$-guided, possibly partial, matches in  $\egame(\psi,\bbS)@(\psi,s)$ in which a position of the form $(\nu q. \psi, r)$ is not visited.

\textsc{Claim.}
% \begin{claim*}\label{p:strategybundledEv}
Let $B^\bbS(\ystrat)$ be the projection of $B(\ystrat)$ on $S$, that is the set
of all paths in $\bbS$ that are a projection on $S$ of a $\ystrat$-guided 
(partial) match in $B(\ystrat)$. Then $B^\bbS(\ystrat)$ is a bundle.
% \end{claim*}
\medskip

\begin{pfclaim}
Assume towards a contradiction that $B^\bbS(\ystrat)$ contains an infinite 
ascending chain $\pi_{0} \sqsubset \pi_{1} \sqsubset \cdots$. 
Let $\pi$ be the limit of this chain and consider the set of elements in 
$B(\ystrat)$ that, projected on $S$, are prefixes of $\pi$. 
By  K\"{o}nig's Lemma, this set contains an infinite ascending chain whose 
limit is an infinite $\ystrat$-guided match in $\egame(\mu p. \phi,\bbS)$
which starts at $(\mu p. \phi,s)$, and of which $\pi$ is the projection on $S$.
By definition of $B(\ystrat)$,  the highest bound variable of $\mu p. \phi$ 
that gets unravelled infinitely often in $\rho$ is a $\mu$-variable, meaning 
that the match is winning for $\abelard$, a contradiction.
\end{pfclaim}

Assume that $s \in \ext{\mu p.\phi}^{\bbS}$, and let $F \isdef  \phi_{p}^{\bbS}$ be the monotone functional defined by $\phi(p)$.
By Proposition \ref{p:unfold=evalgame2}, $\eloise$ has a winning strategy 
$\ystrat'$ in $\egame(\mu p.\phi,\bbS)@(\mu p.\phi,s)$ compatible with a descending winning strategy $\ystrat$ in $\UG_{F}@s$.
By Proposition~\ref{p:afmc-2}, we obtain 
that $s \in \LFP. F_{\rst{T_{\ystrat,s}}}$. 
Because of compatibility, every node in $T_{\ystrat, s}$ occurs on some path of 
$B(\ystrat')$. 
From the Claim
% Claim \ref{p:strategybundledEv}
we known that $B^\bbS(\ystrat')$ is a bundle, meaning that  $T_{\ystrat, s}$ is
noetherian as required. 
\end{proofof}
