% !TEX root = ../00CFVZ_TOCL.tex
\subsection{From automata to formulas}
   \label{sec:parity-to-mc}

It is well-known that there are effective translations from automata to formulas
and vice versa~\cite{ALG02}.
The first result on $\oslang$-automata that we need in this paper is the 
following.

\begin{theorem}\label{t:autofor}
There is an effective procedure that, given an automaton $\bbA$ in 
$\Aut(\oslang)$, returns a formula $\xi_{\bbA} \in \mu\oslang$ which satisfies
the following properties:
\begin{enumerate}[(1)]
    
\item $\xi_{\bbA}$ is equivalent to $\bbA$;

\item $\xi_{\bbA} \in \mu_{N}\oslang$ if $\bbA \in \AutW(\oslang)$;

\item $\xi_{\bbA} \in \mu_{C}\oslang$ if $\bbA \in \AutWC(\oslang)$.
\end{enumerate}
\end{theorem}

In the remainder of this subsection
% (which is not used in other parts of the paper) 
we focus on the proof of this theorem, which is (a refinement of)
a variation of the standard proof showing that any modal automaton can be 
translated into an equivalent formula in the modal $\mu$-calculus (see for
instance~\cite[Section 6]{Ven08}). 
For this reason we will not go into the details of showing that $\bbA$ and 
$\xi_{\bbA}$ are equivalent, but we will provide a detailed definition of the 
translation, and pay special attention to showing that the translations of weak
and of weak-continuous $\oslang$-automata land in the right fragments of 
$\mu\oslang$. 

The definition of $\xi_{\bbA}$ is by induction on the number of clusters  of
$\bbA$ \textcolor{blue}{(see Definition \ref{def:cluster})}, with a subinduction based on the number of states in the top cluster 
of $\bbA$.
For this inner induction we need to widen the class of $\oslang$-automata, and
it will also be convenient to introduce the notion of a preautomaton (which is  
basically an automaton without initital state).

\begin{definition}
A \emph{preautomaton} based on $\oslang$ and $\pprop$, or briefly: a
\emph{preautomaton}, is a triple $\bbA = \tup{A,\tmap,\pmap}$ such that $A$ is
a (possibly empty) finite set of states, $\tmap: A\times \pow(\pprop) \to 
\oslang^+(A)$ and $\pmap: A \to \nat$.

Given a set $X$ of propositional variables, a \emph{generalized preautomaton} 
over $\pprop$ and $X$ is a triple $\bbA = \tup{A,\tmap,\pmap}$ such that $\pmap:
A \to \nat$ is a priority map on the finite state set $A$, while the transition 
map is of the form $\tmap: A\times \pow(\pprop) \to \oslang^+(A\cup X)$.
\end{definition}

Since we will not prove the semantic equivalence of $\bbA$ and $\xi_{\bbA}$,
we confine our attention to the semantics of generalised automata to the
following remark.

\begin{remark}
Generalised automata operate on $\pprop \cup X$-models; it will be convenient to
denote these structures as quintuples of the form $\bbS = \tup{S,R,\tscolors,U,
s_{I}}$, where $\tscolors: \pprop \to \pow S$ is a $\pprop$-colouring and $U: S \to 
\pow X$ is an $X$-valuation on $S$.
The acceptance game $\agame(\bbA,\bbS)$ associated with a generalised automaton
$\bbA = \tup{A,a_{I},\tmap,\pmap}$ and a $\pprop \cup X$-model $\bbS$ is a minor
variation of the one associated with a standard automaton.
At a basic position of the form $(a,s) \in A \times S$, as before $\eloise$ 
needs to come up with a valuation $V$ turning the set $R[s]$ into the domain of 
a one-step model of the formula $\tmap(a,\tscolors(s))$.
The difference with standard automata is that the formula $\tmap(a,\tscolors(s))$
may now involve variables from the set $X$, and that the interpretation of these
is already fixed by the valuation $U$ of $\bbS$, namely by the restriction 
$U_{s}: x \mapsto R[s] \cap U(x)$ to the collection of successors of $s$.
In table, we can present this game as follows:
\begin{center}
\small
\begin{tabular}{|l|c|l|c|} \hline
Position & Player & Admissible moves & Priority \\
\hline
    $(a,s) \in A \times S$
  & $\eloise$
  & $\{V : A \to \pow(R[s]) \mid (R[s],V\cup U_{s}) \models \tmap (a, \tscolors(s)) \}$
  & $\pmap(a)$ 
\\
    $V : A \rightarrow \pow(S)$
  & $\abelard$
  & $\{(b,t) \mid t \in V(b)\}$
  & $0$
\\ \hline
 \end{tabular}
\end{center}
where $V \cup U_{s}$ is the obviously defined $ A \cup X$-valuation on $R[s]$.
\end{remark}

We now turn to the definition of the translation.
We will use the same notation for substitution as for the standard 
$\mu$-calculus, cf.~Subsection~\ref{subsec:mu}.
In addition we use the following notation.

\begin{definition}
Consider, for some preautomaton $\bbA = \tup{A,\tmap,\pmap}$, some state $a \in
A$, and some colour $c \in \pow(\pprop)$, the one-step formula $\tmap(a,c)\in 
\oslang(A)$.
Suppose that for some subset $B \sse A$ we have a collection of 
$\mu\oslang$-formulas $\{ \phi_{b} \mid b \in B \}$.
Without loss of generality we may write $\tmap(a,c) = 
\al(a_{1},\ldots,a_{m},b_{1},\ldots,b_{n})$, where the $a_{i}$ and $b_{j}$ 
belong to $A\setminus B$ and $B$ respectively.
Then we will denote the $\mu\oslang$-formula 
$\nxt{\al}(a_{1},\ldots,a_{m},\phi_{1},\ldots,\phi_{n})$ as follows:
\[
\nxt{\tmap(a,c)}(\phi_{b}/b \mid b \in B)
\isdef \nxt{\al}(a_{1},\ldots,a_{m},\phi_{1},\ldots,\phi_{n}).
\]
\end{definition}

\noindent
We can now define the desired translation from $\oslang$-automata to 
$\mu\oslang$-formulas.


\begin{definition}
\label{d:tr}
By induction on the number of clusters of a preautomaton $\bbA = \tup{A,\tmap,
\pmap}$ we define a map 
\[
\ytr_{\bbA}: A \to \mu\oslang(P).
\]
Based on this definition, for an automaton $\bbA = \tup{A,\tmap,\pmap, a_{I}}$ 
we put
\[
\xi_{\bbA} \isdef \ytr_{\tup{A,\tmap,\pmap}}(a_{I}).
\]

In the base case of the definition of $\ytr$ the preautomaton $\bbA$ has no 
clusters at all, which means in particular that $A = \nada$.
In this case we let $\ytr_{\bbA}$ be the empty map.

In the inductive case we assume that $\bbA = \tup{A,\tmap,\pmap, a_{I}}$ does 
have clusters. 
Let $B \neq \nada$ be the highest cluster \textcolor{blue}{(see Remark \ref{rem:cluster})}, and let $\bbA^{-}$ denote the 
preautomaton with carrier $A \setminus B$, obtained by restricting the maps 
$\tmap$ and $\pmap$ to the set $A \setminus B$.
Then inductively we may assume a translation $\ytr_{\bbA^{-}}: (A \setminus B)
\to \mu\oslang(P)$, and we will define
\[
\ytr_{\bbA}(a) \isdef \ytr_{\bbA^{-}}(a), \quad\text{ if } a \in A \setminus B.
\]

To extend this definition to the states in $B$, we make a case 
distinction.
If $B$ is a degenerate cluster, that is, $B = \{ b \}$ for some state $b$ 
which is not active in itself, then we define
\[
\ytr_{\bbA}(b) \isdef
   \bigvee_{c \in \pow{\pprop}}
   \nxt{\tmap(b,c)}(\ytr_{\bbA^{-}}(a)/a \mid a \in A \setminus B).
\]
The main case of the definition is where $B$ is not degenerate.
Fix an enumeration $b_{1},\ldots,b_{n}$ of $B$ such that $i \leq j$ implies 
$\pmap(b_{i}) \leq \pmap(b_{j})$.
Let $\bbA_{k}$ be the generalized preautomaton\footnote{%
   Here we see the reason to generalise the notion of an automaton:
   in the structure $\bbA_{k}$ ($0 \leq k \leq n$) the objects $b_{k+1},\ldots,
   b_{n}$ are no longer states, but in the formulas $\tmap_{k}(a,c)$ they still 
   occur at the position of states.
}
obtained from $\bbA$ by restricting the transition and priority map to the set
\[
A_{k} \isdef (A\setminus B) \cup \{ b_{1},\ldots,b_{k} \},
\]
so that $\bbA_{0} = \bbA^{-}$ and $\bbA_{n} = \bbA$.
Where $B_{k} \isdef \{ b_{1},\ldots,b_{k} \}$, we now define, by induction on 
$k$, a map 
\[
\ytr^{k}: B \to \mu\oslang(P \cup (B \setminus B_{k})).
\]
In the base case of this definition we set
\[
\ytr^{0}(b) \isdef 
   \bigvee_{c \in \pow{\pprop}} 
   \nxt{\tmap(b,c)}(\ytr_{\bbA^{-}}(a)/a \mid a \in A \setminus B),
\]
and in the inductive case we first define $\eta_{k+1} \isdef \mu$ if
$\pmap(b_{k+1})$ is odd, and $\eta_{k+1} \isdef \nu$ if $\pmap(b_{k+1})$ is 
even, and then set
\[\begin{array}{llll}
     \ytr^{k+1}(b_{k+1}) &\isdef &
   \eta_{k+1} b_{k+1}. \ytr^{0}(b_{k+1})[\ytr^{k}(b_{i})/b_{i} \mid 1 \leq i\leq k]
\\ \ytr^{k+1}(b_{i}) &\isdef &
   \ytr^{k}(b_{i})[\ytr^{k+1}(b_{k+1})/b_{k+1}]
   & \text{ for } i \neq k+1.
\end{array}\]
Finally, we complete the definition of $\ytr_{\bbA}$ by putting
\[
\ytr_{\bbA}(b) \isdef \ytr^{n}(b),
\]
for any $b \in B$.
\end{definition}

In the proof of Theorem~\ref{t:autofor} we will need the following closure 
property of the fragments $\noe{\mu\oslang}{\qprop}$ and 
$\cocont{\mu\oslang}{\qprop}$.

\begin{proposition}
\label{p:comp}
Let $\rprop \subseteq \qprop$ be sets of proposition letters, and let $\phi$ 
and $\phi_{q}$, for each $q \in \qprop$, be formulas in $\mu\oslang$.

(1) If $\phi$ and each $\phi_{q}$ belongs to 
    $\noe{\mu\oslang}{\qprop\setminus\rprop}$ 
    ($\conoe{\mu\oslang}{\qprop\setminus\rprop}$), 
   then so does $\phi[\phi_{q}/q \mid q \in \qprop]$.

(2) If $\phi$ and each $\phi_{q}$ belongs to
   $\cont{\mu\oslang}{\qprop\setminus\rprop}$
   ($\cocont{\mu\oslang}{\qprop\setminus\rprop}$), 
   then so does $\phi[\phi_{q}/q \mid q \in \qprop]$.
\end{proposition}

Both items of this proposition can be proved by a straightforward formula 
induction --- we omit the details.


\begin{proofof}{Theorem~\ref{t:autofor}}
As mentioned, the verification of the equivalence of $\xi_{\bbA}$ and $\bbA$
% for any $\oslang$-automaton $\bbA$, 
is a standard exercise in the theory of parity automata and mu-calculi, and so
we omit the details.
We also skip the proof of item (2), completely focussing on the (harder) third 
item.

To prove this item, it suffices to take an arbitrary continuous-weak 
$\oslang$-preautomaton $\bbA = \tup{A,\tmap,\pmap}$ for the set $\pprop$,
and to show that 
\begin{equation}
\label{eq:tr1}
\ytr_{\bbA}(a) \in \mu_{C}\oslang(\pprop)
\end{equation}
for all $a \in A$.
We will prove this by induction on the number of clusters of $\bbA$.

Since there is nothing to prove in the base case of the proof, we immediately
move to the inductive case.
Let $B$ be the highest cluster of $\bbA$ \textcolor{blue}{(see Remark \ref{rem:cluster})}.
By the induction hypothesis we have $\ytr_{\bbA}(a) = \ytr_{\bbA^{-}}(a) \in 
\mu_{C}\oslang(\pprop)$ for all $a \in A \setminus B$, where 
$\bbA^{-}$ is as in Definition~\ref{d:tr}.
To show that \eqref{eq:tr1} also holds for all $b \in B$, we distinguish cases.

If $B$ is a degenerate cluster, say, $B = \{ b \}$, then for every $c \in 
\pow(\pprop)$, the variables occurring in the formula $\tmap(b,c) \in \oslang(A)$
are all from $A \setminus \{ b \}$.
Given the definition of $\ytr_{\bbA}(b)$, it suffices to show that all formulas 
of the form $\nxt{\tmap(b,c)}(\ytr_{\bbA}(a)/a \mid a \in A \setminus \{ b \})$
belong to the set $\mu_{C}\oslang(\pprop)$, but this is immediate by the 
induction hypothesis and the definition of the language.

If, on the other hand, $B$ is nondegenerate, let $b_{1},\ldots,b_{n}$ enumerate
$B$, and let, for $0\leq k \leq n$, the map $\ytr^{k}: B \to \mu\oslang$ be as in
Definition~\ref{d:tr}.
We only consider the case where $B$ is an odd cluster, i.e., $\pmap(b)$ is odd
for all $b \in B$.
Our key claim here is that 
\begin{equation}
\label{eq:tr2}
\ytr^{k}(b_{i}) \in \mu_{C}\oslang(\pprop \cup \{ b_{k+1},\ldots,b_{n}\}) \cap 
\cont{\mu\oslang}{\{ b_{k+1},\ldots,b_{n}\}},
\end{equation}
for all $k$ and $i$ with $0 \leq k \leq n$ and $0 < i \leq n$.
We will prove this statement by induction on $k$ --- this is the `inner' 
induction that we announced earlier on.

In the base case of this inner induction we need to show that $\ytr^{0}(b_{i})$ 
belongs to both $\mu_{C}\oslang(\pprop \cup B)$ and $\cont{\mu\oslang}{B}$.
Showing the first membership relation is straightforward; for the second, the 
key observation is that by our assumption on $\bbA$, every one-step formula of
the form $\tmap(b_{i},c)$ is syntactically continuous in every variable $b \in 
B$.
Furthermore, by the outer inductive hypothesis we have $\ytr_{\bbA^{-}}(a) \in
\mu_{C}\oslang(\pprop) \subseteq \cont{\mu\oslang}{B}$,
for every $a \in A \setminus B$, and we trivially have that every variable $b 
\in B$ belongs to the set $\cont{\mu\oslang}{B}$.
But then it is immediate by the definition of $\cont{\mu\oslang}{B}$ that 
this fragment contains the formula $\nxt{\tmap(b_{i},c)}(\ytr_{\bbA^{-}}(b)
\mid b \in B)$, and since this fragment is closed under taking disjunctions, 
we find that $\ytr^{0}(b_{i}) \in \cont{\mu\oslang}{B}$ indeed.
This finishes the proof of the base case of the inner induction.

For the inner induction step we fix a $k$ and assume that
% $\ytr^{k}(b_{i}) \in \mu_{C}\oslang(\pprop  \cup \{ b_{k+1},\ldots,b_{n}\}) 
% \cap \cont{\mu\oslang}{\{ b_{k+1},\ldots,b_{n}\}}$,
\eqref{eq:tr2} holds for this $k$ and for all $i$ with $0 < i \leq k$.
We will prove that
\begin{equation}
\label{eq:tr3}
\ytr^{k+1}(b_{i}) \in \mu_{C}\oslang(\pprop) \cap 
    \cont{\mu\oslang}{\{ b_{k+2},\ldots,b_{n}\}}.
\end{equation}
first for $i = k+1$, and then for an arbitrary $i \neq k+1$.
To prove \eqref{eq:tr3} for the case $i = k+1$, first note that
% \begin{equation}
% \label{eq:tr3}
% \ytr^{k+1}(b_{k+1}) \in \mu_{C}\oslang(\pprop) \cap 
%     \cont{\mu\oslang}{\{ b_{k+2},\ldots,b_{n}\}}.
% \end{equation}
% To see this, 
% \[\begin{array}{lcl}
% \ytr^{0}(b_{k+1}) &\in& \mu_{C}\oslang(\pprop\cup \{ b_{k+1},\ldots,b_{n}\}) 
%    \cap \cont{\mu\oslang}{\{ b_{k+1},\ldots,b_{n}\}},
% \end{array}\]
\[\begin{array}{lcl}
\ytr^{0}(b_{k+1}) &\in& \mu_{C}\oslang(\pprop\cup B) 
   \cap \cont{\mu\oslang}{B},
\end{array}\]
as we just saw in the base case of the inner induction.
But then it is immediate by Proposition~\ref{p:comp} and the induction 
hypothesis on the formulas $\ytr^{k}(b_{i})$ that 
\[
\ytr^{0}(b_{k+1})[\ytr^{k}(b_{i})/b_{i} \mid 1 \leq i\leq k]
\in \cont{\mu\oslang}{\{ b_{k+1},\ldots,b_{n}\}},
\]
and from this it easily follows by the definition of 
$\cont{\mu\oslang}{\{ b_{k+1},\ldots,b_{n}\}}$ that 
\[
\mu b_{k+1} b_{k+1}. \ytr^{0}(b_{k+1})[\ytr^{k}(b_{i})/b_{i} \mid 1 \leq i\leq k]
\in \cont{\mu\oslang}{\{ b_{k+2},\ldots,b_{n}\}}
\]
as well. 
That is, 
\[
\ytr^{k+1}(b_{k+1}) \in \cont{\mu\oslang}{\{ b_{k+2},\ldots,b_{n}\}}.
\]
This is the crucial step in proving \eqref{eq:tr3} for the case $i=k+1$,
the proof that $\ytr^{k+1}(b_{k+1}) \in \mu_{C}\oslang(\pprop)$ is easy.

Second, to prove \eqref{eq:tr3} for the case $i \neq k+1$, we first 
% claim that
% \begin{equation}
% \label{eq:tr4}
% \ytr^{k+1}(b_{i}) \in \mu_{C}\oslang(\pprop) \cap 
%     \cont{\mu\oslang}{\{ b_{k+2},\ldots,b_{n}\}}.
% \end{equation}
% To see this,
recall that by the induction hypothesis we have
\[
\ytr^{k}(b_{i}) \in \mu_{C}\oslang(\pprop) \cap 
    \cont{\mu\oslang}{\{ b_{k+1},\ldots,b_{n}\}},
\]
while we just saw that 
$\ytr^{k+1}(b_{k+1}) \in \cont{\mu\oslang}{\{ b_{k+2},\ldots,b_{n}\}}$.
But from the latter two statements it is immediate by Proposition~\ref{p:comp} 
that 
\[
\ytr^{k+1}(b_{i}) = \ytr^{k}(b_{i})[\ytr^{k+1}(b_{k+1})/b_{k+1}] 
\in \cont{\mu\oslang}{\{ b_{k+2},\ldots,b_{n}\}}
\]
so that we have indeed proved \eqref{eq:tr2} for the case $i \neq 
k+1$.
This finishes the proof of the inner induction.

Finally, it follows from \eqref{eq:tr2}, instantiated with $k = n$, 
that for all $b \in B$ we have
\[
\ytr_{\bbA}(b) = \ytr^{n}(b) \in \mu_{C}\oslang(\pprop),
\]
as required for proving the outer induction step.
In other words, we are finished with the proof of \eqref{eq:tr1}, and 
hence, finished with the proof of the theorem.
\end{proofof}


