% !TEX root = ../main.tex

\section{Submodels and quotients}
\label{sec:inv}

There are various natural notions of morphism between monadic models; the one 
that we will be interested here is that  of a (strong) homomorphism.

\begin{definition}
\label{d:hom}
Let $\osmodel = (D,V)$ and $\osmodel' = (D',V')$ be two monadic models.
A map $f: D \to D'$ is a \emph{homomorphism} from $\osmodel$ to $\osmodel'$, 
notation: $f: \osmodel \to \osmodel'$, if we have $d \in V(a)$ iff $f(d) \in 
V'(a)$, for all $a \in A$ and $d \in D$.
\end{definition}

In this section we will be interested in the sentences of $\ofo, \ofoe$ and 
$\ofoei$ that are preserved under taking submodels and the ones that are 
invariant under quotients.

\begin{definition}
\label{d:inv}
Let $\osmodel = (D,V)$ and $\osmodel' = (D',V')$ be two monadic models.
We call $\osmodel$ a \emph{submodel} of $\osmodel'$ if $D \subseteq D'$ and 
the inclusion map $\iota_{DD'}: D \hookrightarrow D'$ is a homomorphism, and 
we say that $\osmodel'$ is a \emph{quotient} of $\osmodel$ if there 
is a surjective homomorphism $f: \osmodel \to \osmodel'$.

Now let $\phi$ be an $\llang$-sentence, where $\llang \in \{ \ofo, \ofoe, \ofoei
\}$.
We say that $\phi$ is \emph{preserved under taking submodels} if 
$\osmodel \models \phi$ implies $\osmodel' \models \phi$, whenever
$\osmodel'$ is a submodel of $\osmodel$.
Similarly, $\phi$ is \emph{invariant under taking quotients} if we have
$\osmodel \models \phi$ iff $\osmodel' \models \phi$, whenever $\osmodel'$ is
a quotient of $\osmodel$.
\end{definition}

The first of these properties (preservation under taking submodels) is well
known from classical model theory --- it is for instance the topic of the
{\L}os-Tarski Theorem.
When it comes to quotients, in model theory one is usually more interested in
the formulas that are \emph{preserved} under surjective homomorphisms (and
the definition of homomorphism may also differ from ours): for instance, this 
is the property that is characterised by Lyndon's Theorem.
Our preference for the notion of \emph{invariance} under quotients stems from 
the fact that the property of invariance under quotients plays a key role in
characterising the \emph{bisimulation-invariant fragments} of various monadic
second-order logics, as is explained in our companion paper~\cite{companionWEAK}.


%%%%
%%%%
%%%%  LOS-TARSKI 
%%%%
%%%%

\subsection{Preservation under submodels}

In this subsection we characterise the fragments of our predicate logics 
consisting of the sentences that are preserved under taking submodels.
That is, the main result of this subsection is a {\L}os-Tarksi Theorem for
$\ofoei$.

\begin{definition}
The \emph{universal fragment} of the set $\ofoei(A)$ is the collection 
$\univ{\ofoei(A)}$ of formulas given by the following grammar:
\[
\varphi \defbnf
\top \mid \bot 
\mid a(x)
\mid \neg a(x)
\mid x \foeq y
\mid x \foneq y
\mid \varphi \lor \varphi
\mid \varphi \land \varphi
\mid \forall x.\varphi
\mid \dqu x.\varphi
\]
where $x,y\in \fovar$ and $a \in A$.
The universal fragment $\univ{\ofoe(A)}$ is obtained by deleting the clause for
$\dqu$ from this grammar, and we obtain the universal fragment $\univ{\ofo(A)}$ 
by further deleting both clauses involving the equality symbol.
\end{definition}

\begin{theorem}
\label{t:univ}
Let $\phi$ be a sentence of the monadic logic $\llang(A)$, where $\llang \in 
\{ \ofo, \ofoe, \ofoei \}$.
Then $\phi$ is preserved under taking submodels if and only if there is a 
equivalent formula $\phi^{\tuniv} \in \univ{\llang(A)}$.
Furthermore, it is decidable whether a sentence $\phi \in \llang(A)$ has this 
property or not.
\end{theorem}

We start by verifying that universal formulas satisfy the property. 

\begin{proposition}
\label{p:univ1}
Let $\phi \in \univ{\llang(A)}$ be a universal sentence of the monadic logic $\llang(A)$, where $\llang \in 
\{ \ofo, \ofoe, \ofoei \}$. 
Then $\phi$ is preserved under taking submodels.
\end{proposition}

\begin{proof} 
It is enough to directly consider the case $\llang = \ofoei$.
Let $(D',V')$ be a submodel of the monadic model $(D,V)$. 
The case for $D = \nada$ being immediate, let us assume $D \neq \nada$.
By induction on the complexity of a formula $\phi \in \univ{\ofoei(A)}$ we will 
show that for any assignment $g: \fovar \to D'$ we have
\[
(D,V),g' \models \phi \text{ implies } (D',V'),g \models \phi,
\]
where $g':= g \circ \iota_{D'D}$.
We will only consider the inductive step of the proof where $\phi$ is of the 
form $\dqu x. \psi$.
Define $X_{D,V} \isdef \{ d \in D \mid (D,V), g'[x \mapsto d] \models \psi \}$,
and similarly, $X_{D',V'} \isdef \{ d \in D' \mid (D',V'), g[x \mapsto d] \models 
\psi \}$.
By the inductive hypothesis we have that $X_{D,V} \cap D' \subseteq X_{D',V'}$,
implying that $D' \setminus X_{D',V'} \subseteq D \setminus X_{D,V}$.
But from this we immediately obtain that 
\[
|D \setminus X_{D,V}| < \omega
\text{ implies } |D' \setminus X_{D',V'}| < \omega,
\]
which means that $(D,V),g' \models \phi$ implies $ (D',V'),g \models \phi$, as 
required. 
\end{proof}

Before verifying the  `hard' side of the theorem, we define the appropriate translations from each monadic logic into its universal fragment.

\begin{definition}
\label{d:univ_trans}
We start by defining the translations for sentences in basic normal forms. 

For  $\ofo$-sentences in basic form we first set
\[
          \Big(  \dbnfofo{\Sigma}  \Big)^{\tuniv} 
         \isdef 
   \forall z \bigvee_{S \in \Sigma} \tau_{S}(z)
\]
Second, we define $(\bigvee_{i} \alpha_{i})^{\tuniv} \isdef \bigvee 
\alpha_{i}^{\tuniv}$. 
Finally, we extend the translation $(-)^{\tuniv}$ to the collection of all 
$\ofo$-sentences by defining $\phi^{\tmoda} \isdef (\tbas{\phi})^{\tuniv}$, 
where $\tbas{\phi}$
is the basic normal form of $\phi$ as given by Fact~\ref{fact:ofonormalform}.

Similarly, for $\ofoe$-sentences we first define
\[
        \Big(  \dbnfofoe{\vlist{T}}{\Pi}  \Big)^{\tuniv} 
         \isdef 
   \forall z \bigvee_{S \in \vlist{T}\cup \Pi} \tau_{S}(z)
\]
and then we extend it to the full language by distributing over disjunction and applying Theorem~\ref{thm:bnfofoe} to convert an arbitrary $\ofoe$ sentence into an equivalent sentence in basic normal form.

Finally, for simple basic formulas of $\ofoei$, the translation $(-)^\tuniv$ is given as follows:
\[
(\dbnfofoei{\vlist{T}}{\Pi}{\Sigma})^{\tuniv} \isdef 
   \forall z \bigvee_{S \in \vlist{T}\cup \Pi\cup\Sigma} \tau_{S}(z) 
   \land \dqu z \bigvee_{S \in \Sigma} \tau_{S}(z).
\]
The definition is thus extended to the full language $\ofoei$ as expected: 
given a $\ofoei$-sentence $\phi$, by Theorem~\ref{thm:bfofoei} and
Proposition~\ref{prop:bfofoei-sigmapi} we compute an equivalent
basic form $\bigvee \dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$, with 
$\Sigma \subseteq \Pi \subseteq \vlist{T}$, and therefore we set
$\phi^{\tuniv}\isdef \bigvee (\dbnfofoei{\vlist{T}}{\Pi}{\Sigma})^{\tuniv}$.
\end{definition}

The missing parts in the proof of the theorem is thence covered by the following
result.

\begin{proposition} 
\label{p:univ2}
For any monadic logic $\llang \in \{ \ofo, \ofoe, \ofoei \}$ there is an 
effective translation $(-)^\tuniv: \llang(A) \to \univ{\llang(A)}$ such that a
sentence $\phi \in \llang(A)$ is preserved under taking submodels if and only
if $\phi\equiv \phi^\tuniv$.
\end{proposition}



\begin{proof}
We only consider the case where $\llang = \ofoei$, leaving the other cases to 
the reader.

It is easy to see that $\phi^{\tuniv} \in \univ{\ofoei(A)}$, for every sentence
$\phi \in \ofoei(A)$; but then it is immediate by Proposition~\ref{p:univ1} that 
$\phi$ is preserved under taking submodels if $\phi \equiv \phi^{\tuniv}$.

For the left-to-right direction, assume that $\phi$ is preserved 
under taking submodels.
It is easy to see that $\phi$ implies $\phi^{\tuniv}$, so we focus on proving
the opposite.
That is, we suppose that $(D,V) \models \phi^{\tuniv}$, and aim to show that 
$(D,V) \models \phi$.
 
By Theorem \ref{thm:bfofoei} and Proposition~\ref{prop:bfofoei-sigmapi} we may assume without loss of 
generality that $\phi$ is a disjunction of setences of the form 
$\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$, where $\Sigma \subseteq \Pi \subseteq 
\vlist{T}$.
It follows that $(D,V)$ satisfies some  disjunct 
$   \forall z \bigvee_{S \in \vlist{T}\cup \Pi\cup\Sigma} \tau_{S}(z) 
   \land \dqu z \bigvee_{S \in \Sigma} \tau_{S}(z)$
of $\Big(\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}\Big)^{\tuniv}$.
Expand $D$ with finitely many elements $\vlist{d}$, in one-one correspondence 
with $\vlist{T}$, and ensure that the type of each $d_{i}$ is $T_{i}$.
In addition, add, for each $S \in \Sigma$, infinitely many elements 
$\{ e^{S}_{n} \mid n \in \omega\}$, each of type $S$.
Call the resulting monadic model $\osmodel' = (D',V')$.

This construction is tailored to ensure that 
$(D',V') \models \dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$, and so we obtain $(D',V') 
\models \phi$.
But obviously, $\osmodel$ is a submodel of $\osmodel'$, whence $(D,V) \models 
\phi$ by our assumption on $\phi$.
\end{proof}

\begin{proofof}{Theorem \ref{t:univ}}
The  first part of  the theorem is an immediate consequence of 
Proposition~\ref{p:univ2}. By applying Fact \ref{f:decido} to 
Proposition~\ref{p:univ2} we finally obtain that for the three concerned 
formalisms the problem of deciding whether a sentence is preserved under taking 
submodels is decidable.
\end{proofof}

As an immediate consequence of the proof of the previous 
Proposition~\ref{p:univ2}, we get effective normal forms for the universal
fragments.

\begin{corollary}\label{cor:univ}
The following hold:
\begin{enumerate}
\item\label{cor:ofo} 
A sentence $\phi \in \ofoe(A)$ is preserved under taking submodels iff it is 
equivalent to a formula 
$\bigvee \big( \forall z \bigvee_{S \in \Sigma} \tau_{S}(z)\big)$, for  types 
$\Sigma\subseteq \wp(A)$.
\item\label{cor:ofoe} 
A sentence $\phi \in \ofoe(A)$ is preserved under taking submodels iff it is 
equivalent to a formula 
$\bigvee \big( \forall z \bigvee_{S \in \vlist{T}\cup \Pi} \tau_{S}(z)\big)$, 
for types $\Pi \subseteq \wp(A)$ and $\vlist{T} \in \wp(A)^k$ for some $k$.
\item\label{cor:ofoei} 
A sentence $\phi \in \ofoei(A)$ is preserved under taking submodels iff it is 
equivalent to a formula 
$\bigvee \big(\forall z \bigvee_{S \in \vlist{T}\cup \Pi\cup\Sigma} \tau_{S}(z) 
   \land \dqu z \bigvee_{S \in \Sigma} \tau_{S}(z)\big)$ 
for types $\Sigma\subseteq \Pi \subseteq \wp(A)$ and $\vlist{T} \in \wp(A)^k$
for some $k$.
\end{enumerate}
In all three cases, normal forms are effective.
\end{corollary}

%%%%
%%%%
%%%%  QUOTIENTS 
%%%%
%%%%

\subsection{Invariance under quotients}

The following theorem states that monadic first-order logic \emph{without}
equality ($\ofo$) provides the quotient-invariant fragment of both monadic 
first-order logic with equality ($\ofoe$), and of infinite-monadic predicate 
logic ($\ofoei$).

\begin{theorem}
\label{t:qinv}
Let $\phi$ be a sentence of the monadic logic $\llang(A)$, where $\llang \in 
\{ \ofoe, \ofoei \}$.
Then $\phi$ is invariant under taking quotients if and only if there is a 
equivalent sentence in $\ofo$.
Furthermore, it is decidable whether a sentence $\phi \in \llang(A)$ has this 
property or not.
\end{theorem}

We first state the `easy' part of the first claim of the theorem. 
Note that in fact, we have already been using this observation in earlier parts
of the paper.

\begin{proposition}
\label{p:m-qinv}
Every sentence in $\ofo$ is invariant under taking quotients.
\end{proposition}

\begin{proof}
Let $f: D \to D'$ provide a surjective homomorphism between the models $(D,V)$ 
and $(D',V')$, and 
observe that for any assignment $g: \fovar \to D$ on $D$, the composition $f 
\circ g: \fovar \to D'$ is an assignment on $D'$.

In order to prove the proposition one may show that, for an arbitrary 
$\ofo$-formula $\phi$ and an arbitrary assignment $g: \fovar \to D$, we have
\begin{equation}
\label{eq:m-qinv1}
(D,V),g \models \phi \text{ iff } (D',V'), f \circ g \models \phi.
\end{equation}
We leave the proof of \eqref{eq:m-qinv1}, which proceeds by a straightforward 
induction on the complexity of $\phi$, as an exercise to the reader.
\end{proof}

To prove the remaining part of Theorem~\ref{t:qinv}, we start with providing 
translations from respectively $\ofoe$ and $\ofoei$ to $\ofo$.

\begin{definition}
\label{d:7-21}
For $\ofoe$-sentences in basic form we first define
\[
   \Big( \dbnfofoe{\vlist{T}}{\Pi} \Big)^{\tmoda} 
\isdef \bigwedge_{i} \exists x_i. \tau_{T_i}(x_i) \land 
\forall x. \bigvee_{S\in\Pi} \tau_S(x),
\]
whereas for $\ofoei$-sentences in basic form we start with defining
\[
\Big( \dbnfofoei{\vlist{T}}{\Pi}{\Sigma} \Big)^{\tmodb} 
\isdef \bigwedge_{i} \exists x_i. \tau_{T_i}(x_i) \land 
\forall x. \bigvee_{S\in\Sigma} \tau_S(x).
\]
In both cases, the translations is then extended to the full language as in Definition \ref{d:univ_trans}.
\end{definition}

Note that the two translations may give \emph{different} translations for 
$\ofoe$-sentences.
Also observe that the $\Pi$ `disappears' in the translation of the formula
$\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$.

The key property of these translations is the following.

\begin{proposition}
\label{p-1P}
\begin{enumerate}
\item
For every one-step model $(D,V)$ and every $\phi \in \ofoe(A)$ we have
\begin{equation}
\label{eq-0cr}
(D,V) \models \phi^{\tmoda} \text{ iff } (D\times \omega,V_\pi) \models \phi.
\end{equation}
\item
For every one-step model $(D,V)$ and every $\phi \in \ofoei(A)$ we have
\begin{equation}
\label{eq-1cr}
(D,V) \models \phi^{\tmodb} \text{ iff } (D\times \omega,V_\pi) \models \phi.
\end{equation}
\end{enumerate}
\noindent
Here $V_{\pi}$ is the induced valuation given by 
$V_{\pi}(a) \isdef \{ (d,k) \mid d \in V(a), k\in\omega\}$.
\end{proposition}

\begin{proof}
We only prove the claim for $\ofoei$ (i.e., the second part of the proposition),
the case for $\ofoe$ being similar.
Clearly it suffices to prove \eqref{eq-1cr} for formulas of the form
$\alpha = \dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$.

First of all, if $\osmodel$ is the empty model, we find ${\vlist{T}}={\Pi} = 
{\Sigma} = \nada$, $(D, V) = (D\times \omega, V_\pi)$, and
$\dbnfofoei{\vlist{T}}{\Pi}{\Sigma} = 
(\dbnfofoei{\vlist{T}}{\Pi}{\Sigma})^\tmodb$. 
In other words, in this case there is nothing to prove.
\smallskip

In the sequel we assume that $D \neq \nada$.

\noindent\fbox{$\Rightarrow$} 
Assume $(D, V) \models \phi^{\tmodb}$, we will show that 
$(D\times \omega, V_\pi) \models \dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$.
Let $d_i$ be such that $V^{\flat}(d_i) = T_{i}$ in $(D, V)$. 
It is clear that the $(d_i,i)$ provide \emph{distinct} elements, with each 
$(d_i,i)$ satisfying $\tau_{T_i}$ in $(D\times\omega, V_{\pi})$ and therefore 
the first-order existential part of $\alpha$ is satisfied. 
With a similar argument it is straightforward to verify that the $\qu$-part of 
$\alpha$ is also satisfied --- here we critically use the observation that
$\Sigma \subseteq \vlist{T}$, so that every type in $\Sigma$ is witnessed in 
the model $(D,V)$, and hence witnessed infinitely many times in $(D\times\omega,
V_\pi)$.

For the universal parts of $\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$ it is enough
to observe that, because of the universal part of $\alpha^\tmodb$, \emph{every}
$d\in D$ realizes a type in $\Sigma$. 
By construction, the same applies to $(D\times\omega, V_{\pi})$, 
therefore this takes care of both universal quantifiers.
\medskip
		
\noindent\fbox{$\Leftarrow$} 
Assuming that $(D\times \omega, V_\pi) \models 
\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$,
we will show that $(D, V) \models \phi^\tmodb$. 
The existential part of $\alpha^{\tmodb}$ is trivial. 
For the universal part we have to show that every element of $D$ realizes a 
type in $\Sigma$. 
Suppose not, and let $d\in D$ be such that $\lnot\tau_S(d)$ for all $S\in 
\Sigma$. 
Then we have $(D\times\omega, V_\pi) \not\models \tau_S(d,k)$ for all $k$.
That is, there are infinitely many elements not realising any type in $\Sigma$. 
Hence we have $(D\times\omega, V_\pi) \not\models \dqu y.\bigvee_{S\in\Sigma} 
\tau_S(y)$. 
Absurd, because this formula is a conjunct of 
$\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$.
\end{proof}

\noindent
We will now show how the theorem follows from this.
First of all we verify that  in both cases $\ofo$ is expressively complete
for the property of being invariant under taking quotients.

\begin{proposition} 
\label{p:invq}
For any monadic logic $\llang \in \{ \ofoe, \ofoei \}$ there is an effective 
translation $(-)^\tinvq: \llang(A) \to \ofo$ such that a sentence $\phi \in
\llang(A)$ is invariant under taking quotients if and only if
$\phi\equiv \phi^\tinvq$.
\end{proposition}
\begin{proof}
Let $\phi$ be a sentence of $\ofoei$, and let $\phi^\tinvq \isdef \phi^{\tmodb}$ (we only cover the case of $\llang = 
\ofoei$, the case for $\llang = \ofoe$ is similar, just take $\phi^\tinvq \isdef \phi^{\tmoda}$)
We will show that 
\begin{equation}
\label{eq:m-qinv2}
\phi \equiv \phi^{\tinvq} \text{ iff } \text{$\phi$ is invariant under 
taking quotients}.
\end{equation}
The direction from right to left is immediate by Proposition~\ref{p:m-qinv}.
For the other direction it suffices to observe that any model $(D,V)$ is a 
quotient of its `$\omega$-product' $(D\times \omega, V_\pi)$, and to 
reason as follows:
\begin{align*}
(D,V) \models \phi 
   & \text{ iff } (D\times \omega, V_\pi) \models \phi
   & \text{(assumption on $\phi$)}
\\ & \text{ iff } (D,V) \models \phi^{\tmodb}
   & \text{(Proposition~\ref{p-1P})}
\end{align*}

\end{proof}

Hence we can conclude.
\begin{proofof}{Theorem~\ref{t:qinv}}
The theorem is an immediate consequence of Proposition \ref{p:invq}.
Finally, the effectiveness of translation  $(\cdot)^{\tmodb}$, decidability of $\ofoei$ (Fact \ref{f:decido}) and \eqref{eq:m-qinv2} yield
that it is decidable whether a given $\ofoei$-sentence $\phi$ is invariant under 
taking quotients or not.
\end{proofof}

As a corollary, we therefore obtain:

\begin{corollary}\label{cor:qinv}
Let $\phi$ be a sentence of the monadic logic $\llang(A)$, where $\llang \in 
\{ \ofoe, \ofoei \}$. Then $\phi$ is invariant under taking quotients if and only if there is a 
equivalent sentence $\bigvee \big(\bigwedge_{S\in\Sigma} \exists x. \tau_S(x) \land 
   \forall x. \bigvee_{S\in\Sigma} \tau_S(x)\big)$, for  types $\Sigma\subseteq \wp(A)$. Moreover, such a normal form is effective.
\end{corollary}

In our companion paper \cite{companionWEAK} on automata, we need versions of these results for 
the monotone and the continuous fragment.
For this purpose we define some slight modifications of the translations 
$(\cdot)^{\tmoda}$ and $(\cdot)^{\tmodb}$ which map positive and syntactically 
continuous sentences to respectively positive and syntactically continuous 
formulas.

\begin{theorem}
\label{t:inv1}
There are effective translations 
$(\cdot)^{\tmoda}: \ofoe^+ \to \ofo^{+} $
 and 
 $(\cdot)^{\tmodb}: {\ofoei}^+ \to \ofo^+$ 
such that 
$\phi \equiv \phi^{\tmoda}$ (respectively, $\phi \equiv \phi^{\tmodb}$) iff 
$\phi$ is invariant under quotients.
Moreover, we may assume that 
$(\cdot)^{\tmodb}: \cont{{\ofoei}(A)}{B}\cap {\ofoei}^+  \to \cont{\ofo(A)}{B} \cap \ofo^+$,
for any $B \subseteq A$.
\end{theorem}
\begin{proof}

We define translations $(\cdot)^{\tmoda}:\ofoe^+ \to\ofo^+$ and 
$(\cdot)^{\tmodb}: {\ofoei}^+ \to\ofo^+$ as follows. 
For $\ofoe^{+},{\ofoei}^{+}$-sentences in simple basic form we define
\[\begin{array}{lll}
     \Big( \posdbnfofoe{\vlist{T}}{\Pi} \Big)^{\tmoda} 
   & \isdef 
   & \bigwedge_{i} \exists x_i. \tau^+_{T_i}(x_i) 
     \land \forall x. \bigvee_{S\in\Pi} \tau^+_S(x),
\\   \Big( \posdbnfofoei{\vlist{T}}{\Pi}{\Sigma} \Big)^{\tmodb} 
   & \isdef 
   & \bigwedge_{i} \exists x_i. \tau^+_{T_i}(x_i) 
     \land \forall x. \bigvee_{S\in\Sigma} \tau^+_S(x),
\end{array}\]
and then we use, respectively, the Corollaries~\ref{cor:ofoepositivenf} and
\ref{cor:ofoeipositivenf} to extend these translations to the full positive
fragments $\ofoe^{+}$ and ${\ofoei}^{+}$, as we did in Definition~\ref{d:7-21}
for the full language.

We leave it as an exercise for the reader to prove the analogue of 
Proposition~\ref{p-1P} for these translations, and to show how the first
statements of the theorem follows from this.

Finally, to see why we may assume that $(\cdot)^{\tmodb}$ restricts to a map 
from the syntactically $B$-continuous fragment of ${\ofoei}^+(A)$ to the 
syntactically $B$-continuous fragment of ${\ofo}^+(A)$, assume that $\phi 
\in \ofoei(A)$ is continuous in $B \subseteq A$.
By Corollary~\ref{cor:ofoeicontinuousnf} we may assume that $\phi$ is a
disjunction of formulas of the form $\posdbnfofoei{\vlist{T}}{\Pi}{\Sigma}$,
where $B \cap \bigcup \Sigma = \nada$.
This implies that in the formula $\phi^{\tmodb}$ no predicate symbol $b \in
B$ occurs in the scope of a universal quantifier, and so $\phi^{\tmodb}$
is syntactically continuous in $B$ indeed.
\end{proof}

