% !TEX root = ../main.tex

\section{Continuity}
\label{sec-cont}

In this section we study the sentences that are \emph{continuous} in some set 
$B$ of monadic predicate symbols.

\begin{definition}\label{def:cont}
Let $U$ and $V$ be two $A$-valuations on the same domain $D$.
For a set $B \subseteq A$, we write $U \leq^{\omega}_{B} V$ if $U \leq_{B} V$ 
and $U(b)$ is finite, for every $b \in B$.

Given a monadic logic $\llang$ and a formula $\phi \in \llang(A)$ we say that
$\phi$ is \emph{continuous in $B \subseteq A$} if $\phi$ is monotone in $B$
and satisfies the following:
\begin{equation}
\label{eq:cont}
\text{if } (D, V), g \models \phi \text{ then } 
(D,  U), g \models \phi \text{ for some } U \leq^{\omega}_{B} V.
\end{equation}
for every monadic model $(D,V)$ and every assignment $g:\fovar\to D$.
\end{definition}

\begin{remark}\label{rem:contprodeach}
As for monotonicity, but with slightly more effort, one may show that a formula 
$\phi$ is continuous in a set $B$ iff it is continuous in every $b \in B$.
\end{remark}

What explains both the name and the importance of this property is its 
equivalence to so called \emph{Scott continuity}. 
To understand it, we may formalise the dependence of the meaning of a  monadic 
sentence $\phi$ with $m$-free variables $\vlist{x}$ in a one-step model $\osmodel
= (D,V )$ on a fixed name $b \in A$ as a map $\phi^\osmodel_b : \wp(D) \to 
\wp(D^m)$ defined by 
\[
X \subseteq D \mapsto \{ \vlist{d} \in D^m \mid  
(D,V[b \mapsto X]) \models \phi(\vlist{d}) \}.
\]
One can then verify that a sentence $\phi$ is continuous in $b$ if and only if 
the operation  $\phi^\osmodel_b$ is continuous with respect to the Scott 
topology on the powerset algebras\footnote{%
   The interested reader is referred to \cite[Sec. 8]{FV12} for a more precise 
   discussion of the connection.}. 
Scott continuity is of key importance in many areas of theoretical computer 
sciences where ordered structures play a role, such as domain theory (see 
e.g. \cite{abramsky1994domain}).

Similarly as for monotonicity, the semantic property of continuity can also be 
provided with a corresponding syntactical formulation.

\begin{definition}
Let $\llang \in \{ \ofo, \ofoe \}$
The fragment of $\ofo(A)$ of formulas that are 
\emph{syntactically continuous} in a subset $B \subseteq A$ is 
defined by the following grammar:
\[
\phi \defbnf \psi 
   \mid b(x) 
   \mid \phi \land \phi \mid \phi \lor \phi
   \mid \exists x.\phi,
\]
where $b\in B$ and $\psi \in \llang(A\setminus B)$. 
In both cases, we let $\cont{\llang(A)}{B}$ denote the set of $B$-continuous 
sentences.
\end{definition}

To define the syntactically continuous fragment of $\ofoei$, we first introduce
the following binary generalised quantifier $\wqu$: given two formulas $\phi(x)$ 
and $\psi$, we set
\[
\wqu x.(\phi,\psi) \isdef \forall x.(\phi(x) \lor \psi(x)) \land \dqu x.\psi(x).
\]
The intuition behind $\wqu$ is the following. If $(D,V),g \models \wqu x.(\phi,
\psi)$, then because of the second conjunct there are only finitely many $d \in
D$ refuting $\psi$. 
The point is that this weakens the universal quantification of the first conjunct
to the effect that only the finitely many mentioned elements refuting $\psi$ need
to satisfy $\phi$.

\begin{definition}
The fragment of $\ofoei(A)$-formulas that are 
\emph{syntactically continuous} in a subset $B \subseteq A$ is given by the 
following grammar:
\[
\phi \defbnf \psi 
   \mid a(x) 
   \mid \phi \land \phi \mid \phi \lor \phi 
   \mid \exists x.\phi \mid \wqu x.(\phi,\psi),
\]
where $b\in B$ and $\psi \in \ofoei(A\setminus B)$. 
We let $\cont{\ofoei(A)}{B}$ denote the set of $B$-continuous $\ofoei$-sentences.
\end{definition}


For $\ofo$ and $\ofoe$, the equivalence between the semantical and syntactical 
properties of continuity was established by van Benthem 
in \cite{van1997dynamic}. 

\begin{proposition}\label{fact:vb}
Let $\phi$ be a sentence of the monadic logic $\llang(A)$, where $\llang \in 
\{ \ofo, \ofoe\}$.
Then $\phi$ is continuous in a set $B \subseteq A$ if and only if there is a 
equivalent sentence $\phi^{\tcont} \in \cont{\llang(A)}{B}$.
\end{proposition}

\begin{proof}
The direction from right to left is covered by Proposition \ref{p:coniscont} 
below, so we immediately turn to the completeness part of the statement.
The case of $\ofo$ being treated in Subsection \ref{subsec:conofo}, we only discuss  the statement for $\ofoe$.
Hence, let $\phi \in \ofoe(A)$ be continuous in 
$B$. 
For simplicity in the exposition, we assume $B=\{b\}$, the case of an arbitrary 
$B$ being easily generalisable from what follows. 
Let $\vlist{y} \isdef y_0 \dots y_{k-1}$ be a list of $k$ variables not
occurring in $\phi$. 
Consider the formula $\phi_k( \vlist{y} )$ obtained from $\phi$  by substituting 
each occurrence of an atomic formula of the form $b(x)$ with the formula 
$\bigvee_{\ell < k} x = y_\ell$.
Define $\Phi \isdef \{ \exists \vlist{y}.\phi_k( \vlist{y} ) \mid k \in \omega \} \cup \{ \phi_{\emodel}\}$, 
where $ \phi_{\emodel} \isdef \forall x. \bot$ if $\emodel \models \phi$ and 
$ \phi_{\emodel} \isdef \exists x. \bot$ otherwise. 
By construction $\Phi \subset \cont{\ofoe(A)}{B}$.
Now, notice that $\lnot \Phi \cup \{\phi\}$ is inconsistent. 
Hence, by compactness of first-order logic, there is a $k \in \omega$ such that
$\phi \models \bigvee_{\ell < k} \exists \vlist{y}.\phi_k( \vlist{y} ) \lor 
  \phi_{\emodel}$. 
By monotonicity, $\exists \vlist{y}.\phi_k( \vlist{y} ) \models \phi$, for every
$k \in \omega$, and by definition $\phi_{\emodel} \models \phi$.
We therefore conclude that $\phi \equiv \bigvee_{\ell < k} \exists \vlist{y}.
\phi_k( \vlist{y} ) \lor \phi_\emodel$. 
As $\cont{\ofoe(A)}{B}$ is closed under disjunctions, this ends the proof of 
the statement.
\end{proof}

In this paper, we extend such a characterisation to $\ofoei$. 
Moreover, analogously to what we did in the previous section, for $\ofo$ and 
$\ofoei$ we provide both an explicit translation and a decidability result. 
From this latter perspective, the case of $\ofoe$ remains however open.

\begin{theorem}
\label{t:cont}
Let $\phi$ be a sentence of the monadic logic $\llang(A)$, where $\llang \in 
\{ \ofo, \ofoei \}$.
Then $\phi$ is continuous in a set $B \subseteq A$ if and only if there is a 
equivalent sentence $\phi^{\tcont} \in \cont{\llang(A)}{B}$.
Furthermore, it is decidable whether a sentence $\phi \in \llang(A)$ has this 
property or not.
\end{theorem}

Analogously to the previous case of monotonicity, the proof of the theorem is 
composed of two parts.  
We start with the right-left implication of the first claim (the preservation
statement), which also holds for $\ofoe$.

\begin{proposition}\label{p:coniscont}
Every sentence $\phi \in \cont{\llang(A)}{B}$ is continuous in $B$,
where $\llang\in\{ \ofo, \ofoe, \ofoei \}$.
\end{proposition}

\begin{proof}
First observe that $\phi$ is monotone in $B$ by 
Proposition~\ref{p:monoismonot}.
The case for $D=\nada$ being clear, we assume $D\neq\nada$.
We show, by induction, that any one-step formula $\phi$ in the fragment (which
may not be a sentence) satisfies \eqref{eq:cont}, for every non-empty one-step
model $(D, V)$ and assignment ${ g:\fovar\to D}$.
%
\begin{enumerate}[\textbullet]
\item 
If $\phi = \psi \in \llang(A\setminus B)$, changes in the $B$ part of the 
valuation will not affect the truth value of $\phi$ and hence the condition is 
trivial. 

\item
Case $\phi = b(x)$ for some $b \in B$: if $(D, V), g \models b(x)$ then 
$g(x)\in  V(b)$. 
Let $U$ be the valuation given by $U(b) \isdef \{ g(x) \}$,
$U(a) \isdef \nada$ for $a \in B \setminus \{b \}$ and
$U(a) \isdef V(a)$ for $a \in A \setminus B$.
Then it is obvious that $(D,  U), g \models b(x)$, while it is immediate by the
definitions that $U \leq^{\omega}_{B} V$.

\item 
Case $\phi = \phi_1 \lor \phi_2$: assume $(D, V), g \models \phi$. 
Without loss of generality we can assume that $(D, V), g \models \phi_1$ and 
hence by induction hypothesis there is $U \leq^{\omega}_{B} V$ such that $(D, U),
g \models \phi_1$ which clearly implies $(D, U), g \models 
\phi$. 

\item 
Case $\phi = \phi_1 \land \phi_2$: assume $(D, V), g \models \phi$. 
By induction hypothesis we have $U_1,U_2 \leq^{\omega}_{B} V$ such that 
$(D,U_{1}), g \models \phi_1$ and $(D, U_2), g \models \phi_2$.
Let $U$ be the valuation defined by putting $U(a) \isdef U_{1}(a) \cup U_{2}$; 
then clearly we have $U \leq^{\omega}_{B} V$, while it follows by monotonicity
that $(D,U), g \models \phi_1$ and $(D, U), g \models \phi_2$.
Clearly then $(D, U), g \models \phi$.

\item
Case $\phi = \exists x.\phi'(x)$ and $(D, V), g \models \phi$. 
By definition there exists $d\in D$ such that $(D, V), g[x\mapsto d] \models 
\phi'(x)$. 
By induction hypothesis there is a valuation $U \leq^{\omega}_{B} V$ such that 
$(D, U), g[x\mapsto d] \models \phi'(x)$ and hence $(D, U), g \models 
\exists x.\phi'(x)$.
\item
Case $\phi = \wqu x.(\phi',\psi)\in \cont{\ofoei(A)}{B}$ and $(D, V), g \models \phi$.  Define the formulas $\alpha(x)$ and 
$\beta$ as follows:
\[
\phi = 
\forall x.\underbrace{(\phi'(x) \lor \psi(x))}_{\alpha(x)} \land 
\underbrace{\dqu x.\psi(x)}_\beta.
\]

Suppose that $(D, V), g \models \phi$. 
By the induction hypothesis, for every $d \in D$ which  satisfies $(D, V), g_d
\models \alpha(x)$ (where we write $ g_d \isdef g[x\mapsto d]$) there is a 
valuation $U_d \leq^{\omega}_{B} V$ such that $(D, U_d), g_d \models \alpha(x)$. 
The crucial observation is that because of $\beta$, only finitely many elements
of $d$ refute $\psi(x)$. 
Let $U$ be the valuation defined by putting $U(a) \isdef \bigcup \{U_d(a) \mid 
(D, V), g_d \not\models \psi(x) \}$. 
Note that for each $b \in B$, the set $U(b)$ is a finite union of finite sets, 
and hence finite itself; it follows that $U \leq^{\omega}_{B} V$.
We claim that
\begin{equation}
\label{eq:1801}
(D, U), g \models \phi.
\end{equation}
%
It is clear that $(D, U), g \models \beta$ because $\psi$ (and hence $\beta$) 
is $B$-free.
To prove that $(D, U), g \models \forall x\, \alpha(x)$, take an arbitrary $d 
\in D$, then we have to show that $(D, U), g_d \models \phi'(x) \lor \psi(x)$. 
We consider two cases: If $(D, V), g_d \models \psi(x)$ we are done, again
because $\psi$ is $B$-free.
On the other hand, if $(D, V), g_d \not\models \psi(x)$, then $(D, U_d), g_d 
\models \alpha(x)$ by assumption on $U_{d}$, while it is obvious that $U_{d} 
\leq_{B} U$; but then it follows by monotonicity of $\alpha$ that $(D, U), g_d
\models \alpha(x)$.

\end{enumerate}
This finishes the proof.
\end{proof}


The second part of the proof of the theorem, is thus constituted by the following stronger version of the expressive completeness result that provides as a corollary normal forms for the syntactically continuous fragments.
\begin{proposition}
\label{p:efftranscont}
Let $\llang$ be one of the logics $\{ \ofo, \ofoei \}$.
There exists an effective translation $(-)^\tcont:\llang(A) \to \cont{\llang(A)}{B}$ such that
a sentence ${\phi \in \llang(A)}$ is continuous in $B \subseteq A$ if and only if 
$\phi\equiv \phi^\tcont$.
\end{proposition}

We prove the two manifestations of Proposition \ref{p:efftranscont} separately,
in two respective subsections.

By putting together the two propositions above, we are thence able to conclude.
%\begin{proofof}
\begin{proof}[Proof of Theorem \ref{t:cont}]
The first claim follows from  Proposition~\ref{p:efftranscont}. 
Hence, by applying Fact \ref{f:decido} to Proposition \ref{p:efftranscont}, the 
problem of checking whether a sentence $\phi \in \llang(A)$ is continuous in 
$B \subseteq A$ or not, is decidable.
\end{proof}

We conjecture that Proposition \ref{p:efftranscont}, and therefore
Theorem \ref{t:cont}, holds also for $\llang=\ofoe$.

%%%%%
%%%%%  FO
%%%%%
\subsection{Continuous fragment of $\ofo$}\label{subsec:conofo}

Since continuity implies monotonicity, by Theorem \ref{t:mono}, in order to 
verify the $\ofo$-variant of Proposition \ref{p:efftranscont}, it is enough to
proof the following result.

\begin{proposition}\label{prop:ofocont}
There is an effective translation $(-)^\tcont:\monot{\ofo(A)}{B} \to 
\cont{\ofo(A)}{B}$ such that a sentence $\phi \in \monot{\ofo(A)}{B}$ is
continuous in $B \subseteq A$ if and only if $\phi\equiv \phi^\tcont$.
\end{proposition}

\begin{proof}
By Corollary \ref{cor:ofopositivenf}, to define the translation we assume, without loss of generality, that $\phi$
is in the basic form $\bigvee \mondbnfofo{\Sigma}{B}$.
For the translation, let
\[
(\bigvee \mondbnfofo{\Sigma}{B})^\tcont \isdef 
\bigvee \mondgbnfofo{\Sigma}{\Sigma^{-}_{B}}{B}
\]
where $\Sigma^{-}_{B} \isdef \{S\in \Sigma \mid B \cap S = \nada \}$.
From the construction it is clear that $\phi^\tcont \in \cont{\ofo(A)}{B}$ and
therefore the right-to-left direction of the proposition is immediate by 
Proposition~\ref{p:coniscont}. 

For the left-to-right direction assume that $\phi$ is continuous in $B$, we have
to prove that $(D, V) \models \phi$ iff $(D, V) \models \phi^\tcont$, for every 
one-step model $(D, V)$.
Our proof strategy consists of proving the same equivalence for the model 
$(D\times \omega, V_\pi)$, where $D\times\omega$ consists of $\omega$ many 
copies of each element in $D$ and $V_\pi$ is the valuation given by $V_{\pi}(a) 
\isdef \{(d,k) \mid d\in  V(a), k\in\omega\}$.
It is easy to see that $(D, V) \equiv^{\ofo} (D\times \omega, V_\pi)$ (see 
Proposition~\ref{p-1P}) and so it suffices indeed to prove that
\[
(D\times\omega, V_\pi) \models \phi
\text{ iff }
(D\times\omega, V_\pi) \models \phi^\tcont.
\]
Consider first $D= \nada$.
Then $(D\times\omega, V_\pi) = \emodel$, and therefore the claim is true since 
$\mondbnfofo{\nada}{B} = \mondgbnfofo{\nada}{\nada^{-}_{B}}{B}$ and
$\emodel \models  \mondbnfofo{\Sigma}{B}$ iff $\Sigma=\nada$. 
Hence, assume $D \neq \nada$.
\bigskip

\noindent \fbox{$\Rightarrow$}
Let $(D\times\omega, V_\pi) \models \phi$.
As $\phi$ is continuous in $B$ there is a valuation $U \leq^{\omega}_{B} V_\pi$ 
satisfying $(D\times\omega, U) \models \phi$. 
This means that  $(D\times\omega, U) \models \mondbnfofo{\Sigma}{B}$ for some
disjunct $\mondbnfofo{\Sigma}{B}$ of $\phi$.
Below we will use the following fact (which can easily be verified):
\begin{equation}
\label{eq:con101}
(D\times\omega),U \models \tau^{B}_{S}(d,k) \text{ iff }
S \setminus B = U^{\flat}(d,k) \setminus B \text{ and }
S \cap B \subseteq U^{\flat}(d,k).
\end{equation}

Our claim is now that $(D\times\omega, U) \models 
\mondgbnfofo{\Sigma}{\Sigma^{-}_{B}}{B}$. 

The existential part of $\mondgbnfofo{\Sigma}{\Sigma^{-}_{B}}{B}$ is trivially 
true. 
To cover the universal part, it remains to show that every element of 
$(D\times\omega, U)$ realizes a $B$-positive type in $\Sigma^{-}_{B}$.
Take an arbitrary pair $(d,k) \in D\times\omega$ and let $T$ be the (full) type 
of $(d,k)$, that is, let $T \isdef U^{\flat}(d,k)$.
If $B \cap T = \nada$ then trivially $T\in \Sigma^{-}_{B}$ and we are done. 
So suppose $B \cap T \neq \nada$. 
Observe that in $D\times\omega$ we have infinitely many copies of $d\in D$.
Hence, as $U(b)$ is finite for every $b \in B$, there must be some $(d,k')$ 
with type $U^{\flat}(d,k') = V_{\pi}^{\flat}(d,k') \setminus B = 
V_{\pi}^{\flat}(d,k) \setminus B = T \setminus B$.
It follows from $(D\times\omega, U) \models \mondbnfofo{\Sigma}{B}$ and 
\eqref{eq:con101} that there is some $S \in \Sigma$ such that
$S \setminus B = U^{\flat}(d,k') \setminus B = U^{\flat}(d,k')$ and 
$S \cap B \subseteq U^{\flat}(d,k) \cap B = \nada$.
From this we easily derive that $S = U^{\flat}(d,k')$ and $S \in \Sigma^{-}_{B}$.
Finally, we observe that $S \setminus B = U^{\flat}(d,k') \setminus B =
U^{\flat}(d,k) \setminus B$ and $S \cap B = \nada \subseteq U^{\flat}(d,k)$, 
so that by \eqref{eq:con101} we find that $D \times \omega,U) \models 
\tau^{B}_{S}(d,k)$ indeed.

Finally, by monotonicity it directly follows from
$(D\times\omega, U) \models \mondgbnfofo{\Sigma}{\Sigma^{-}_{B}}{B}$
that 
$(D\times\omega, V_{\pi}) \models \mondgbnfofo{\Sigma}{\Sigma^{-}_{B}}{B}$,
and from this it is immediate that $(D\times\omega, V_\pi) \models \phi^\tcont$.
\bigskip

\noindent \fbox{$\Leftarrow$}
Let $(D\times\omega, V_\pi) \models 
\mondgbnfofo{\Sigma}{\Sigma^{-}_{B}}{B}$.
To show that $(D\times\omega, V_\pi) \models \mondbnfofo{\Sigma}{B}$, the 
existential part is trivial. For the universal part just observe that 
$\Sigma^{-}_{B} \subseteq \Sigma$.
\end{proof}

A careful analysis of the translation gives us the following corollary,
providing normal forms for the continuous fragment of $\ofo$.

\begin{corollary}\label{cor:ofocontinuousnf}
For any sentence $\phi \in \ofo(A)$, the following hold.
\begin{enumerate}
\item 
The formula $\phi$ is continuous in $B \subseteq A$ iff it is equivalent to a 
formula $\bigvee \mondgbnfofo{\Sigma}{\Sigma^{-}_{B}}{B}$ for some types $\Sigma
\subseteq \wp(A)$, where $\Sigma^{-}_{B} \isdef \{S\in \Sigma \mid B \cap S = 
\nada \}$.
\item 
If $\phi$ is monotone in $A$ 
then $\phi$ is continuous in $B \subseteq A$ iff it is equivalent to a formula 
in the basic form $\bigvee \posdgbnfofo{\Sigma}{\Sigma^{-}_{B}}$ for some types 
$\Sigma \subseteq \wp(A)$, where $\Sigma^{-}_{B} \isdef \{S\in \Sigma \mid B \cap
S = \nada \}$.
\end{enumerate}
\end{corollary}

%%%%%
%%%%%  FOEI
%%%%%


\subsection{Continuous fragment of $\ofoei$}

As for the previous case, the $\ofoei$-variant of
Proposition \ref{p:efftranscont} is an immediate consequence of 
Theorem \ref{t:mono} and the following proposition.

\begin{proposition}\label{lem:ofoeictrans}
There is an effective translation $(-)^\tcont:\monot{\ofoei(A)}{B} \to 
\cont{\ofoei(A)}{B}$ such that a sentence $\phi \in \monot{\ofoei(A)}{B}$ is 
continuous in $B$ if and only if $\phi\equiv \phi^\tcont$.
\end{proposition}

\begin{proof}
By Corollary \ref{cor:ofoeipositivenf}, we assume that $\phi$ is in basic normal form, i.e., $\phi = \bigvee 
\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}$.
For the translation let 
$\big(\bigvee \mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}\big)^\tcont \isdef 
\bigvee \mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}^\tcont$ where
\[
\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}^\tcont \isdef
\begin{cases}
   \bot &\text{ if } B \cap \bigcup \Sigma \neq \nada
\\ \mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B} &\text{ otherwise}.
\end{cases}
\]

First we prove the right-to-left direction of the proposition. 
By Proposition~\ref{p:coniscont} it is enough to show that $\phi^\tcont \in
\cont{\ofoei(A)}{B}$. 
We focus on the disjuncts of $\phi^\tcont$. 
The interesting case is where $B \cap \bigcup \Sigma = \nada$.
If we rearrange $\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}$ somewhat and define
the formulas $\phi', \psi$ as follows:
%
\begin{align*}
\exists \vlist{x}.\Big(
   & \arediff{\vlist{x}} \land \bigwedge_i \tau^B_{T_i}(x_i)\ 
     \land \forall z.(\underbrace{\lnot\arediff{\vlist{x},z} 
         \lor \bigvee_{S\in \Pi} \tau^B_S(z)}_{\phi'(\vlist{x},z)} 
         \lor \underbrace{\bigvee_{S\in \Sigma} \tau^B_S(z)}_{\psi(z)})\ 
     \land \dqu y.\underbrace{\bigvee_{S\in\Sigma} \tau^B_S(y)}_{\psi(y)} \Big)
\\ &  
     \land \bigwedge_{S\in\Sigma} \qu y.\tau^B_S(y).
\end{align*}
%
Then we find that
\[
\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B} \equiv 
\exists \vlist{x}.\Big(\arediff{\vlist{x}} 
     \land \bigwedge_i \tau^B_{T_i}(x_i) \land \wqu z.(\phi'(\vlist{x},z),\psi(z)) 
      \Big) 
   \land \bigwedge_{S\in\Sigma} \qu y.\tau^B_S(y),
\]
which belongs to the required fragment because $B \cap \bigcup \Sigma = \nada$.

For the left-to-right direction of the proposition we have to prove that $\phi \equiv
\phi^\tcont$.

\bigskip
\noindent\fbox{$\Rightarrow$} 
Let $(D, V) \models \phi$. 
Because $\phi$ is continuous in $B$ we may assume that $ V(b)$ is finite, for
all $b \in B$.
Let $\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}$ be a disjunct of $\phi$ such that
$(D, V) \models \mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}$.
If $D = \nada$, then $ {\vlist{T}}={\Pi}={\Sigma}=\nada$, and 
$\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B} =
(\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B})^\tcont$. 
Hence, let $D \neq \nada$.
%
Suppose for contradiction that $B \cap \bigcup \Sigma \neq \nada$, then there 
must be some $S\in\Sigma$ with $B \cap S \neq \nada$. 
Because $(D, V) \models \mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}$ we have, in
particular, that $(D, V) \models \qu y.\tau^B_S(x)$ and hence $V(b)$ must be 
infinite, for any $b \in B \cap S$, which is absurd.
It follows that $B \cap \bigcup \Sigma = \nada$, but then we trivially conclude
that $(D, V) \models \phi^\tcont$ because the disjunct remains unchanged. 

\bigskip
\noindent\fbox{$\Leftarrow$} 
Let $(D, V) \models \phi^\tcont$. 
The only difference between $\phi$ and $\phi^\tcont$ is that some disjuncts may 
have been replaced by $\bot$. Therefore this direction is trivial.
\end{proof}

We conclude the section by stating the following corollary, 
providing normal forms for the continuous fragment of $\ofoei$. 

\begin{corollary}\label{cor:ofoeicontinuousnf}
For any sentence $\phi \in \ofoei(A)$, the following hold.
\begin{enumerate}
\item \label{pt:ofoeifcontinuous}
The formula $\phi$ is continuous in $B \subseteq A$ iff 
$\phi$ is equivalent to a 
formula, effectively obtainable from $\phi$, which is a disjunction of 
formulas $\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}$
where $\vlist{T},\Sigma$ and $\Pi$ are such that $\Sigma \subseteq \Pi \subseteq 
\vlist{T}$ and $B \cap \bigcup\Sigma = \nada$. 	
\item \label{pt:ofoeimonotone}
If $\phi$ is monotone 
(i.e., $\phi\in{\ofoei}^+(A)$) then 
$\phi$ is continuous in $B \subseteq A$
iff
it is equivalent to a formula, effectively obtainable from $\phi$, which is a 
disjunction of formulas
$\bigvee \posdbnfofoei{\vlist{T}}{\Pi}{\Sigma}$,
where $\vlist{T},\Sigma$ and $\Pi$ are such that $\Sigma \subseteq \Pi \subseteq 
\vlist{T}$ and $B \cap \bigcup\Sigma = \nada$. 	
\end{enumerate}
\end{corollary}

\begin{proof}
Notice that, from Proposition \ref{prop:bfofoei-sigmapi}, every sentence in the
basic form $\bigvee \dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$
can be assumed such that $\Sigma \subseteq 
\Pi \subseteq \vlist{T}$. The claims hence follow by construction of the translation.
\end{proof}

