% !TEX root = ../main.tex

\section{Monotonicity}
\label{sec-mono}

In this section we provide our first characterisation result, which concerns the
notion of monotonicity. 

\begin{definition}\label{def:mono}
Let $V$ and $V'$ be two valuations on the same domain $D$, then we say that $V'$
is a \emph{$B$-extension of} $V$, notation: $V \leq_{B} V'$, if $V(b) \subseteq
V'(b)$ for every $b \in B$, and $V(a) = V'(a)$ for every $a \in A \setminus B$.

Given a monadic logic $\llang$ and a formula $\phi \in \llang(A)$ we say 
that $\phi$ is \emph{monotone in $B \subseteq A$} if 
\begin{equation}
\label{eq:mono}
(D,V),g \models \phi \text{ and } V \leq_{B} V' \text{ imply } 
(D,V'),g \models \phi,
\end{equation}
for every pair of monadic models $(D,V)$ and $(D,V')$ and every assignment 
$g:\fovar\to D$.
\end{definition}

\begin{remark}\label{rem:monotprodeach}
It is easy to prove that a formula is monotone in $B \subseteq A$ if and only if 
it is monotone in every $b \in B$. 
\end{remark}

The semantic property of monotonicity can usually be linked to the syntactic 
notion of positivity.
Indeed, for many logics, a formula $\phi$ is monotone in $a \in A$ iff 
$\phi$ is equivalent to a formula where all occurrences of $a$ have a positive
polarity, that is, they are situated in the scope of an even number of 
negations.

\begin{definition}
For $\llang \in \{ \ofo, \ofoe \}$ we define the fragment of $A$-formulas that 
are \emph{positive} in all predicates in $B$, in short: the \emph{$B$-positive 
formulas} by the following grammar:
\[
\phi \defbnf  \psi \mid b(x) 
  \mid \phi \land \phi \mid \phi \lor \phi
  \mid \exists x.\phi \mid \forall x.\phi, 
\]
where $b \in B$ and $\psi \in \llang(A\setminus B)$ (that is, there are no 
occurrences of any $b \in B$ in $\psi$).
Similarly, the $B$-positive fragment of $\ofoei$ is given by
\[
\phi \defbnf  \psi \mid b(x) 
  \mid \phi \land \phi \mid \phi \lor \phi
  \mid \exists x.\phi \mid \forall x.\phi 
  \mid \qu x.\phi \mid \dqu x.\phi, 
\]
where $b\in B$ and $\psi \in \ofoei(A\setminus B)$. 

In all three cases, we let $\monot{\llang(A)}{B}$ denote the set of $B$-positive
sentences.
\end{definition}

Note that the difference between the fragments $\monot{\ofo(A)}{B}$ and 
$\monot{\ofoe(A)}{B}$ lies in the fact that in the latter case, the `$B$-free'
formulas $\psi$ may contain the equality symbol. 
Clearly $\monot{\llang(A)}{A}= \llang^+$.

\begin{theorem}
\label{t:mono}
Let $\phi$ be a sentence of the monadic logic $\llang(A)$, where $\llang \in 
\{ \ofo, \ofoe, \ofoei \}$.
Then $\phi$ is monotone in a set $B \subseteq A$ if and only if there is a 
equivalent formula $\phi^{\tmono} \in \monot{\llang(A)}{B}$.
Furthermore, it is decidable whether a sentence $\phi \in \llang(A)$ has this 
property or not.
\end{theorem}

The `easy' direction of the first claim of the theorem is taken care of by the following 
proposition.

\begin{proposition}
\label{p:monoismonot}
Every formula $\phi \in \monot{\llang(A)}{B}$ is monotone in $B$,
where $\llang$ is one of the logics $\{ \ofo, \ofoe, \ofoei \}$.
\end{proposition}

\begin{proof} 
The case for $D= \nada$ being immediate, we assume $D \neq \nada$.
The proof is a routine argument by induction on the complexity of $\phi$.
That is, we show by induction, that any formula $\phi$ in the $B$-positive
fragment (which may not be a sentence) satisfies \eqref{eq:mono}, for every 
monadic model $(D,V)$, valuation $V' \geq_{B} V$ and assignment ${g:\fovar\to 
D}$.
We focus on the generalised quantifiers. 
Let $(D,V),g \models \phi$ and
$V \leq_{B} V'$.
%
\begin{enumerate}[\textbullet]
\item
Case $\phi = \qu x.\phi'(x)$. By definition there exists an infinite set
$I\subseteq D$ such that for all $d\in I$ we have $(D,V),g[x\mapsto d] \models 
\phi'(x)$. 
By induction hypothesis $(D,V'),g[x\mapsto d] \models\phi'(x)$ 
for all $d \in I$. Therefore $(D,V'),g \models \qu x.\phi'(x)$.

\item Case $\phi = \dqu x.\phi'(x)$. 
Hence there exists $C\subseteq D$ such that for all $d\in C$ we have 
$(D,V),g[x\mapsto d] \models \phi'(x)$ and $D\setminus C$ is \emph{finite}. 
By induction hypothesis $(D,V'),g[x\mapsto d] \models \phi'(x)$ for all $d 
\in C$. 
Therefore $(D,V'),g \models \dqu x.\phi'(x)$.
\end{enumerate}
%
This finishes the proof.
\end{proof}

The `hard' direction of the first claim of the theorem states that the fragment
$\monot{\ofo}{B}$ is complete for monotonicity in $B$. In order to prove it,
we need to show that every sentence which is monotone in $B$ is equivalent to 
some formula in $\monot{\ofo}{B}$. 
We actually are going to prove a stronger result.

\begin{proposition}
\label{p:efftrans}
Let $\llang$ be one of the logics $\{ \ofo, \ofoe, \ofoei \}$.
There exists an effective translation $(-)^\tmono:\llang(A) \to \monot{\llang(A)}{B}$ such that
a sentence ${\phi \in \llang(A)}$ is monotone in $B \subseteq A$ %if and 
only if 
$\phi\equiv \phi^\tmono$.
\end{proposition}

We prove the three manifestations of Proposition \ref{p:efftrans} separately,
in three respective subsections.

%\begin{proofof}
\begin{proof}[Proof of Theorem \ref{t:mono}]
The first claim of the Theorem is an immediate consequence of
Proposition~\ref{p:efftrans}.
By effectiveness of the translation and Fact \ref{f:decido}, it is therefore 
decidable whether a sentence $\phi \in \llang(A)$ is monotone in $B \subseteq A$
or not.
%\end{proofof}
\end{proof}
The following definition will be used throughout in the remaining of the section.

\begin{definition}
Given $S \subseteq A$ and $B \subseteq A$ we use the following notation
\[
\tau^{B}_S(x) \isdef  \bigwedge_{b\in S} b(x) \land 
   \bigwedge_{b\in A\setminus (S\cup B)}\lnot b(x) ,
\]
for what we call the \emph{$B$-positive} $A$-type $\tau^{B}_S$.
\end{definition}

Intuitively, $\tau^{B}_S$ works almost like the $A$-type $\tau_S$, the 
difference being that $\tau^{B}_S$ discards the negative information for the 
names in $B$.
If $B = \{a\}$ we write $\tau^a_S$ instead of $\tau^{\{a\}}_S$. 
Observe that with this notation, $\tau^+_S$ is equivalent to $\tau^A_S$.

\subsection{Monotone fragment of $\ofo$}

In this subsection we prove the $\ofo$-variant of Proposition~\ref{p:efftrans}.
That is, we give a translation that constructively maps arbitrary sentences
into $\monot{\ofo}{B}$ and that moreover it preserves truth iff the given
sentence is monotone in $B$.
To formulate the translation we need to introduce some new notation. 

\begin{definition}\label{def:monbasicformofoe}
Let $B\subseteq A$ be a finite set of names. 
The $B$-positive variant of $\dbnfofo{\Sigma}$ is given as follows:
\[
\mondbnfofo{\Sigma}{B} \isdef  
\bigwedge_{S\in\Sigma} \exists x. \tau^{B}_S(x) \land 
  \forall x. \bigvee_{S\in\Sigma} \tau^{B}_S(x).
\]
We also introduce the following generalised forms of the above notation:
\[
\mondgbnfofo{\Sigma}{\Pi}{B} \isdef  
\bigwedge_{S\in\Sigma} \exists x. \tau^{B}_S(x) \land 
  \forall x. \bigvee_{S\in\Pi} \tau^{B}_S(x).
\]
The \emph{positive} variants of the above notations are defined as 
$\posdbnfofo{\Sigma} \isdef  \mondbnfofo{\Sigma}{A}$ and 
$\posdgbnfofo{\Sigma}{\Pi} \isdef  \mondgbnfofo{\Sigma}{\Pi}{A}$.
\end{definition}


\begin{proposition}
\label{p:fomon}
There exists an effective translation $(-)^\tmono:\ofo(A) \to \monot{\ofo(A)}{B}$ such that
a sentence ${\phi \in \ofo(A)}$ is monotone in $B\subseteq A$ if and only if 
$\phi\equiv \phi^\tmono$.
\end{proposition}
%
\begin{proof}
To define the translation, by Fact \ref{fact:ofonormalform}, we assume, without loss of generality, that $\phi$
is in the normal form $\bigvee \dbnfofo{\Sigma}$ given in 
Definition~\ref{def:bfofo}, where
$\dbnfofo{\Sigma} = 
\bigwedge_{S\in\Sigma} \exists x. \tau_S(x) \land 
  \forall x. \bigvee_{S\in\Sigma} \tau_S(x)$.
We define the translation as
\[
(\bigvee \dbnfofo{\Sigma})^\tmono\isdef  \bigvee \mondbnfofo{\Sigma}{B}.
\]
From the construction it is clear that $\phi^\tmono \in \monot{\ofo(A)}{B}$ 
and therefore the right-to-left direction of the proposition is immediate by 
Proposition~\ref{p:monoismonot}. 
For the left-to-right direction assume that $\phi$ is monotone in $B$, we 
have to prove that $(D,V) \models \phi$ if and only if $(D,V) \models 
\phi^\tmono$.

\bigskip
\noindent \fbox{$\Rightarrow$} This direction is trivial.

\bigskip
\noindent \fbox{$\Leftarrow$} Assume $(D,V) \models \phi^\tmono$ and let 
$\Sigma$ be such that $(D,V) \models \mondbnfofo{\Sigma}{B}$. 
If $D = \nada$, then $\Sigma = \nada$ and $\mondbnfofo{\Sigma}{B}= 
\dbnfofo{\Sigma}$. 
Hence, assume $D \neq \nada$, and clearly $\Sigma \neq \nada$.

Because of the existential part of $\mondbnfofo{\Sigma}{B}$, every type $S \in
\Sigma$ has a `$B$-witness' in $\osmodel$, that is, an element $d_{S} \in D$
such that $(D,V) \models \tau^{B}_{S}(d_{S})$.
It is in fact safe to assume that all these witnesses are \emph{distinct}
(this is because $(D,V)$ can be proved to be $\ofo$-equivalent to such a model,
cf.~Proposition~\ref{p-1P}).
But because of the universal part of $\mondbnfofo{\Sigma}{B}$, we may assume
that for all states $d$ in $D$ there is a type $S_{d}$ in $\Sigma$ such that
$(D,V) \models \tau^{B}_{S_{d}}(d)$.
Putting these observations together we may assume that the map $d \mapsto S_{d}$
is surjective.

Note however, that where we have $(D,V) \models \tau^{B}_{S}(d)$, this does not 
necessarily imply that $(D,V) \models \tau_{S}(d)$: it might well be the case
that $d \in V(b)$ but $b \not\in S_{d}$, for some $ b \in B$.
What we want to do now is to shrink $V$ in such a way that the witnessed type
($S_d$) and the actually satisfied type coincide.
That is, we consider the valuation $U$ defined as $U^{\flat}(d) \isdef 
S_d$.\footnote{%
   Recall that a valuation $U:A\to\wp(D)$ can also be represented as a 
   colouring $U^{\flat}: D\to\wp(A)$ given by $U^{\flat}(d) \isdef 
   \{a \in A \mid d\in V(a)\}$.
   } 
It is then immediate by the surjectivity of the map $d \mapsto S_{d}$ that 
$(D,U) \models \dbnfofo{\Sigma}$, which implies that $(D,U) \models \phi$.

We now claim that 
\begin{equation}
\label{eq:moninf1}
U \leq_{B} V.
\end{equation}
To see this, observe that for $a \in A \setminus B$ we have the following 
equivalences:
\[
d \in U(a) \iff a \in S_{d} \iff (D,V) \models a(d) \iff d \in V(a),
\]
while for $b \in B$ we can prove
\[
d \in U(b) \iff b \in S_{d} \Longrightarrow (D,V) \models b(d) \iff d \in V(b).
\]
This suffices to prove \eqref{eq:moninf1}.

But from \eqref{eq:moninf1} and the earlier observation that $(D,U) \models 
\phi$ it is immediate by the monotonicity of $\phi$ in $B$ that $(D,V) \models 
\phi$.
\end{proof}

A careful analysis of the translation gives us the following 
corollary, providing normal forms for the monotone fragment of $\ofo$.

\begin{corollary}\label{cor:ofopositivenf}
For any sentence $\phi \in \ofo(A)$, the following hold.
\begin{enumerate}
\item 
The formula $\phi$ is monotone in $B \subseteq A$ iff it is equivalent to a 
formula in the basic form $\bigvee \mondbnfofo{\Sigma}{B}$ for some types 
$\Sigma \subseteq \wp(A)$.
\item The formula $\phi$ is monotone in every $a\in A$ 
iff $\phi$ is equivalent to a formula $\bigvee \posdbnfofo{\Sigma}$ for some
types $\Sigma \subseteq \wp(A)$.
\end{enumerate}
In both cases the norma forms are effective.
\end{corollary}

\subsection{Monotone fragment of $\ofoe$}

In order to prove the $\ofoe$-variant of Proposition~\ref{p:efftrans}, we need
to introduce some new notation. 

\begin{definition}
Let $B\subseteq A$ be a finite set of names. 
The $B$-monotone variant of $\dbnfofoe{\vlist{T}}{\Pi}$ is given as follows:
\begin{align*}
\mondbnfofoe{\vlist{T}}{\Pi}{B} 
  &\isdef  \exists \vlist{x}.\big(\arediff{\vlist{x}} 
     \land \bigwedge_i \tau^{B}_{T_i}(x_i) 
     \land \forall z.(\arediff{\vlist{x},z} \to \bigvee_{S\in \Pi} \tau^{B}_S(z))
     \big). 	
\end{align*}
When the set $B$ is a singleton $\{a\}$ we will write $a$ instead of $B$. 
The positive variant $\posdbnfofoe{\vlist{T}}{\Pi}$ of 
$\dbnfofoe{\vlist{T}}{\Pi}$ is defined as above but with $+$ in place of $B$.
\end{definition}

\begin{proposition}
\label{p:monofoeismonot}
There exists an effective translation $(-)^\tmono:\ofoe(A) \to \monot{\ofoe(A)}{B}$ such
that a sentence ${\phi \in \ofoe(A)}$ is monotone in $B$ if and only if 
$\phi\equiv \phi^\tmono$.
\end{proposition}

\begin{proof}
In proposition~\ref{p:mono-ofoei} this result is proved for $\ofoei$ (i.e.,
$\ofoe$ extended with generalised quantifiers). 
It is not difficult to adapt the proof for $\ofoe$. 
The translation is defined as follows. By Theorem \ref{thm:bnfofoe} ,without loss of generality, assume that $\phi$ is in basic normal form $\bigvee \dbnfofoe{\vlist{T}}{\Pi}$. Then 
 $\phi^\tmono \isdef \bigvee \mondbnfofoe{\vlist{T}}{\Pi}{B}$.
\end{proof}


Combining the normal form for $\ofoe$ and the proof of the above proposition, we 
therefore obtain a normal form for the monotone fragment of
$\ofoe$.

\begin{corollary}\label{cor:ofoepositivenf}
For any sentence $\phi \in \ofo(A)$, the following hold.
\begin{enumerate}
\item The formula $\phi$ is monotone in $B\subseteq A$ iff it is equivalent
to a formula in the basic form $\bigvee \mondbnfofoe{\vlist{T}}{\Pi}{B}$ where
for each disjunct we have $\vlist{T} \in \wp(A)^k$ for some $k$ and 
$\Pi\subseteq\vlist{T}$.
		
\item 
The formula $\phi$ is monotone in all $a\in A$ iff it is equivalent to a formula
in the basic form 
$\bigvee \posdbnfofoe{\vlist{T}}{\Pi}$ where for each disjunct we have 
$\vlist{T} \in \wp(A)^k$ for some $k$ and $\Pi\subseteq\vlist{T}$.
\end{enumerate}
In both cases, normal forms are effective.
\end{corollary}

\subsection{Monotone fragment of $\ofoei$}

First, in this case too we introduce some notation for the positive variant of
a sentence in normal form.

\begin{definition}
Let $B\subseteq A$ be a finite set of names. 
The $B$-positive variant of $\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$ is given as 
follows:
\begin{align*}
    \mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B} 
  & \isdef  \mondbnfofoe{\vlist{T}}{\Pi \cup \Sigma}{B} 
       \land \mondbnfinf{\Sigma}{B}
\\  \mondbnfofoe{\vlist{T}}{\Lambda}{B} 
  & \isdef  \exists \vlist{x}.\big(\arediff{\vlist{x}} 
        \land \bigwedge_i \tau^{B}_{T_i}(x_i) 
	\land \forall z.(\arediff{\vlist{x},z} 
	    \to \bigvee_{S\in\Lambda} \tau^{B}_S(z))
	\big)
\\ \mondbnfinf{\Sigma}{B} 
  &\isdef  \bigwedge_{S\in\Sigma} \qu y.\tau^{B}_S(y) 
       \land \dqu y.\bigvee_{S\in\Sigma} \tau^{B}_S(y).
\end{align*}
When the set $B$ is a singleton $\{a\}$ we will write $a$ instead of $B$. 
The positive variant of $\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$ is defined as 
$\posdbnfofoei{\vlist{T}}{\Pi}{\Sigma} \isdef  
\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{A}$.
\end{definition}

\noindent
We are now ready to proceed with the proof of the $\ofoei$-variant of
Proposition~\ref{p:efftrans} and thus to give the translation.

\begin{proposition}
\label{p:mono-ofoei}
There is an effective translation $(-)^\tmono:\ofoei(A) \to \monot{\ofoei(A)}{B}$
such that a sentence ${\phi \in \ofoei(A)}$ is monotone in $B$ if and only if
$\phi\equiv \phi^\tmono$.
\end{proposition}
\begin{proof}
By Theorem \ref{thm:bfofoei}, we assume that $\phi$ is in the normal form
$\bigvee\dbnfofoei{\vlist{T}}{\Pi}{\Sigma} = 
\dbnfofoe{\vlist{T}}{\Pi \cup \Sigma} \land \dbnfinf{\Sigma}$
for some sets of types $\Pi,\Sigma \subseteq \wp(A)$ and each $T_i \subseteq A$.
For the translation we define
\[
\Big(\bigvee \dbnfofoei{\vlist{T}}{\Pi}{\Sigma}\Big)^\tmono\isdef  
\bigvee \mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}.
\]
%
From the construction it is clear that $\phi^\tmono \in \monot{\ofoei(A)}{B}$
and therefore the right-to-left direction of the proposition is immediate by 
Proposition~\ref{p:monoismonot}. 
For the left-to-right direction assume that $\phi$ is monotone in $B$, we 
have to prove that $(D,V) \models \phi$ if and only if $(D,V) \models 
\phi^\tmono$.

\bigskip
\noindent \fbox{$\Rightarrow$} This direction is trivial.

\bigskip
\noindent \fbox{$\Leftarrow$}
Assume $(D,V) \models \phi^\tmono$, and in particular that $(D,V) \models 
\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}$. 
If $D = \nada$, then $\Sigma = \Pi = \vlist{T} = \nada$ and 
$\mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}= \dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$. 
Hence, assume $D \neq \nada$.
Observe that the elements of $D$ can be partitioned in the following way:
%
\begin{enumerate}[(a)]
\itemsep 0 pt
\item distinct elements $t_i \in D$ such that each $t_i$ satisfies 
   $\tau^{B}_{T_i}(x)$,
\item\label{it:dpi} for every $S \in \Sigma$ an infinite set $D_S$,
   such that every $d \in D_S$ satisfies $\tau^{B}_{S}$,
\item\label{it:ds} 
a \emph{finite} set $D_\Pi$ of elements, each satisfying one of the $B$-positive
   types $\tau^{B}_{S}$ with $S \in \Pi$.
\end{enumerate}
%
Following this partition, with every element $d\in D$ we may associate a type
$S_{d}$ in, respectively, (a)~$\vlist{T}$, (b)~$\Sigma$, or (c)~$\Pi$, such 
that $d$ satisfies $\tau^{B}_{S_{d}}$.
As in the proof of proposition~\ref{p:fomon}, we now consider the valuation $U$ 
defined as $U^{\flat}(d) \isdef S_d$, and as before we can show that
$U \leq_{B} V$.
Finally, it  easily from the definitions that $(D,U) \models 
\dbnfofoei{\vlist{T}}{\Pi}{\Sigma}$, implying that $(D,U) \models \phi$.
But then by the assumed $B$-monotonicity of $\phi$ it is immediate that $(D,V) 
\models \phi$, as required.
\end{proof}


As with the previous two cases, the translation provides normal forms for the monotone fragment of $\ofoei$.

\begin{corollary}\label{cor:ofoeipositivenf}
For any sentence $\phi \in \ofoei(A)$, the following hold:
\begin{enumerate}
\item 
The formula $\phi$ is monotone in $B \subseteq A$ iff it is equivalent to
a formula $\bigvee \mondbnfofoei{\vlist{T}}{\Pi}{\Sigma}{B}$ for 
$\Sigma\subseteq\Pi \subseteq \wp(A)$ and $\vlist{T} \in \wp(A)^k$ for some $k$.
\item 
The formula $\phi$ is monotone in every $a\in A$ iff it is equivalent to a
formula in the basic form $\bigvee \posdbnfofoei{\vlist{T}}{\Pi}{\Sigma}$ for 
types $\Sigma\subseteq \Pi \subseteq \wp(A)$ and $\vlist{T} \in \wp(A)^k$ for 
some $k$.
\end{enumerate}
In both cases, normal forms are effective.
\end{corollary}

\begin{proof}
We only remark that to obtain $\Sigma\subseteq\Pi$ in the above normal forms 
it is enough to use Proposition~\ref{prop:bfofoei-sigmapi} before applying 
the translation.
\end{proof}


